\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red},
  numbers=left,
  numberstyle=\tiny\color{gray},
  breaklines=true,
  frame=single,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 2001 -- Score}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Statement Summary}
Given a decimal number as a string of $N$ digits and a budget of $K$
adjacent swaps, find the maximum and minimum numbers obtainable.
Leading zeros are not permitted (except for the number 0 itself).

\section{Solution: Greedy Selection Sort}

\subsection{Maximizing}
Process positions left to right. At position $i$, locate the largest
digit in the range $[i, \min(i+K, N-1)]$ (breaking ties by choosing
the leftmost occurrence to minimize swaps). Bubble it to position $i$
by swapping with its left neighbor repeatedly, deducting the number of
swaps used from $K$.

\begin{lemma}
The greedy strategy is optimal for maximization.
\end{lemma}
\begin{proof}
At each step we place the largest reachable digit into the
leftmost unfilled position. Any other choice yields a lexicographically
smaller result, because the first position where the two results differ
will have a smaller digit in the non-greedy solution.
\end{proof}

\subsection{Minimizing}
The same strategy applies, but we select the \emph{smallest} digit.
For position 0, we skip digit `0' to avoid leading zeros---we select
the smallest \emph{non-zero} digit reachable within $K$ swaps.

\subsection{Edge Case}
If the string is all zeros (representing the number 0), no leading-zero
avoidance is needed.

\section{C++ Implementation}

\begin{lstlisting}
#include <bits/stdc++.h>
using namespace std;

string maximize(string s, int K) {
    int n = s.size();
    for (int i = 0; i < n - 1 && K > 0; i++) {
        int bestPos = i;
        for (int j = i + 1; j < n && j - i <= K; j++)
            if (s[j] > s[bestPos])
                bestPos = j;
        for (int j = bestPos; j > i; j--) {
            swap(s[j], s[j - 1]);
            K--;
        }
    }
    return s;
}

string minimize(string s, int K) {
    int n = s.size();
    for (int i = 0; i < n - 1 && K > 0; i++) {
        int bestPos = i;
        for (int j = i + 1; j < n && j - i <= K; j++) {
            if (i == 0) {
                // Avoid leading zero
                if (s[j] == '0') continue;
                if (s[bestPos] == '0' || s[j] < s[bestPos])
                    bestPos = j;
            } else {
                if (s[j] < s[bestPos])
                    bestPos = j;
            }
        }
        // If all reachable digits for position 0 are '0', skip
        if (i == 0 && s[bestPos] == '0') continue;
        for (int j = bestPos; j > i; j--) {
            swap(s[j], s[j - 1]);
            K--;
        }
    }
    return s;
}

int main() {
    ios::sync_with_stdio(false);
    cin.tie(nullptr);

    string S;
    int K;
    cin >> S >> K;

    cout << minimize(S, K) << "\n";
    cout << maximize(S, K) << "\n";

    return 0;
}
\end{lstlisting}

\section{Complexity Analysis}
\begin{itemize}
  \item \textbf{Time:} $O(N \cdot \min(N, K))$. For each of $N$
        positions, we scan at most $\min(K, N)$ positions ahead.
        The total number of swaps across all iterations is at most $K$.
  \item \textbf{Space:} $O(N)$.
\end{itemize}

\end{document}
