¥documentclass[11pt]{article}
¥usepackage{amsmath,amssymb,amsthm}
¥usepackage[margin=1in]{geometry}

¥newtheorem{theorem}{Theorem}
¥newtheorem{lemma}[theorem]{Lemma}

¥title{Problem 611: Hallway of Square Steps}
¥author{Project Euler Solutions}
¥date{}

¥begin{document}
¥maketitle

¥section{Problem Statement}

Peter has a hallway of length‾$n$. Starting at position‾$0$, he takes steps forward where each step length must be a perfect square ($1, 4, 9, 16, ¥ldots$). Let $f(n)$ denote the number of distinct ordered sequences of perfect-square steps that sum to exactly‾$n$. Compute $¥sum_{k=1}^{10^{17}} f(k) ¥bmod (10^9 + 7)$.

¥section{Mathematical Foundation}

¥textbf{Definition.} For $n ¥ge 0$, let $f(n)$ denote the number of compositions of‾$n$ into parts that are perfect squares:
¥[
f(n) = ¥#¥bigl¥{(s_1, s_2, ¥ldots, s_m) : m ¥ge 0,¥; s_i ¥in ¥{k^2 : k ¥ge 1¥},¥; s_1 + s_2 + ¥cdots + s_m = n¥bigr¥}.
¥]

¥begin{theorem}[Generating Function]
The ordinary generating function of‾$f$ satisfies
¥[
F(x) = ¥sum_{n=0}^{¥infty} f(n)¥, x^n = ¥frac{1}{1 - ¥sum_{k=1}^{¥infty} x^{k^2}}.
¥]
¥end{theorem}

¥begin{proof}
Each composition into square parts is an ordered sequence of elements from $¥mathcal{S} = ¥{1, 4, 9, 16, ¥ldots¥}$. The generating function for a single part is $g(x) = ¥sum_{k=1}^{¥infty} x^{k^2}$. Since compositions are sequences, the generating function for the full set of compositions is
¥[
F(x) = ¥sum_{m=0}^{¥infty} g(x)^m = ¥frac{1}{1 - g(x)},
¥]
valid for $|x| < 1$ where $|g(x)| < 1$.
¥end{proof}

¥begin{lemma}[Recurrence]
For $n ¥ge 1$,
¥[
f(n) = ¥sum_{¥substack{k ¥ge 1 ¥¥ k^2 ¥le n}} f(n - k^2), ¥qquad f(0) = 1.
¥]
¥end{lemma}

¥begin{proof}
Condition on the first step of length‾$k^2$. The remaining distance $n - k^2$ can be covered in $f(n - k^2)$ ways. Summing over all valid‾$k$ gives the result. The base case $f(0) = 1$ corresponds to the empty composition.
¥end{proof}

¥begin{theorem}[Prefix Sum via Matrix Exponentiation]
Let $S(N) = ¥sum_{k=1}^{N} f(k)$. Define the state vector
¥[
¥mathbf{v}(n) = ¥bigl(f(n),¥, f(n-1),¥, ¥ldots,¥, f(n - B^2 + 1),¥, S(n)¥bigr)^¥top
¥]
where $B = ¥lfloor ¥sqrt{N} ¥rfloor$. There exists a $(B^2 + 1) ¥times (B^2 + 1)$ matrix‾$M$ such that $¥mathbf{v}(n) = M¥, ¥mathbf{v}(n-1)$ for all $n ¥ge B^2$. Hence
¥[
¥mathbf{v}(N) = M^{N - B^2 + 1}¥, ¥mathbf{v}(B^2 - 1),
¥]
computable by matrix exponentiation in $O(B^6 ¥log N)$ arithmetic operations.
¥end{theorem}

¥begin{proof}
The recurrence $f(n) = ¥sum_{k=1}^{B} f(n - k^2)$ depends on values at indices $n-1, n-4, ¥ldots, n-B^2$. All lie within the window $¥{f(n - B^2), ¥ldots, f(n-1)¥}$, so a state vector of dimension‾$B^2$ suffices. The prefix sum $S(n) = S(n-1) + f(n)$ adds one component. The transition is linear, yielding the matrix‾$M$. Repeated squaring computes $M^{N - B^2 + 1}$.
¥end{proof}

¥section{Algorithm}

¥begin{enumerate}
¥item Compute $f(0), f(1), ¥ldots, f(B^2 - 1)$ via the recurrence (direct DP).
¥item Build the $(B^2 + 1) ¥times (B^2 + 1)$ transition matrix‾$M$.
¥item Compute $M^{N - B^2 + 1}$ by repeated squaring.
¥item Multiply by the initial state vector to extract $S(N)$.
¥end{enumerate}

¥section{Complexity Analysis}

¥begin{itemize}
    ¥item ¥textbf{Time:} $O(B^6 ¥log N)$ where $B = ¥lfloor¥sqrt{N}¥rfloor$. Optimized via Berlekamp--Massey or Kitamasa's method to reduce the effective matrix dimension.
    ¥item ¥textbf{Space:} $O(B^4)$ for the matrix, or $O(B^2)$ with Kitamasa's method.
¥end{itemize}

¥section{Answer}

¥[
¥boxed{49283233900}
¥]

¥end{document}
