\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red!70!black},
  numbers=left,
  numberstyle=\tiny\color{gray},
  breaklines=true,
  frame=single,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 2011 -- Ricehub}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Summary}
There are $R$ rice fields at sorted positions $x_1 < x_2 < \cdots < x_R$ along a line. A hub can be placed at any integer position. The cost of connecting field~$i$ to a hub at position~$h$ is $|x_i - h|$. Given budget~$B$, find the maximum number of fields connectable to a single hub.

\section{Solution}

\subsection{Key Observations}

\begin{lemma}
The optimal hub for a contiguous set of fields $\{x_l, \ldots, x_r\}$ is at the median $x_m$ where $m = \lfloor(l+r)/2\rfloor$.
\end{lemma}
\begin{proof}
The sum of absolute deviations $\sum_{i=l}^{r} |x_i - h|$ is minimized when $h$ is the median of the set, a standard result.
\end{proof}

\begin{lemma}
The optimal set of fields is a contiguous subarray of the sorted positions.
\end{lemma}
\begin{proof}
If two fields $x_a$ and $x_b$ with $x_a < x_b$ are connected, then every field between them has cost $\le \max(|x_a - h|, |x_b - h|)$, so including it cannot increase the cost.
\end{proof}

\subsection{Algorithm: Two Pointers with Prefix Sums}
Use a sliding window $[l, r]$. For each right endpoint $r$, expand while the cost is within budget $B$, and shrink $l$ when it exceeds $B$. The cost for window $[l, r]$ with median at $m = \lfloor(l+r)/2\rfloor$:
\[
\text{cost} = x_m(m - l + 1) - S_m + S_r - x_m(r - m)
\]
where $S_k = \sum_{i=0}^{k} x_i$ is the prefix sum (using appropriate indexing).

\subsection{Complexity}
\begin{itemize}
    \item \textbf{Time:} $O(R)$ with two pointers.
    \item \textbf{Space:} $O(R)$.
\end{itemize}

\section{Implementation}

\begin{lstlisting}
#include <bits/stdc++.h>
using namespace std;

int besthub(int R, int L, int X[], long long B) {
    vector<long long> prefix(R + 1, 0);
    for (int i = 0; i < R; i++)
        prefix[i + 1] = prefix[i] + X[i];

    auto cost = [&](int l, int r) -> long long {
        int m = (l + r) / 2;
        long long leftCost = (long long)X[m] * (m - l + 1)
                           - (prefix[m + 1] - prefix[l]);
        long long rightCost = (prefix[r + 1] - prefix[m])
                            - (long long)X[m] * (r - m + 1);
        return leftCost + rightCost;
    };

    int ans = 0, l = 0;
    for (int r = 0; r < R; r++) {
        while (cost(l, r) > B)
            l++;
        ans = max(ans, r - l + 1);
    }

    return ans;
}
\end{lstlisting}

\end{document}
