\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red!70!black},
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=8pt,
  frame=single,
  breaklines=true,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 1995 -- Packing Bits}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

Given a sequence of bits (0s and 1s) represented in a specific input format,
pack them into bytes (groups of 8 bits) and output the result.

\medskip
\noindent\textbf{Input format:}
\begin{itemize}
  \item First line: $n$, the total number of bits.
  \item Following lines: bits given as integers (each either 0 or 1),
        up to 8 values per line.
\end{itemize}

\noindent\textbf{Output format:}
Pack every 8 consecutive bits into one byte (most significant bit first) and output
each byte as a decimal integer (0--255), with 8 values per output line.
If the last group has fewer than 8 bits, pad with zeros on the right.

\section{Solution Approach}

This is a straightforward simulation problem.

\begin{enumerate}
  \item Read all $n$ bits into an array.
  \item Pad the array with zeros so that its length is a multiple of 8.
  \item Process the bits in groups of 8. For each group, compute the byte value:
        \[
          \text{byte} = \sum_{i=0}^{7} b_i \cdot 2^{7-i}.
        \]
  \item Output the resulting bytes, 8 per line.
\end{enumerate}

\section{C++ Solution}

\begin{lstlisting}
#include <cstdio>
#include <vector>
using namespace std;

int main() {
    int n;
    scanf("%d", &n);

    vector<int> bits(n);
    for (int i = 0; i < n; i++)
        scanf("%d", &bits[i]);

    // Pad to a multiple of 8
    while (bits.size() % 8 != 0)
        bits.push_back(0);

    int numBytes = bits.size() / 8;
    for (int i = 0; i < numBytes; i++) {
        int val = 0;
        for (int j = 0; j < 8; j++)
            val = val * 2 + bits[i * 8 + j];

        if (i > 0 && i % 8 == 0)
            printf("\n");
        else if (i > 0)
            printf(" ");
        printf("%d", val);
    }
    printf("\n");

    return 0;
}
\end{lstlisting}

\section{Complexity Analysis}

\begin{itemize}
  \item \textbf{Time complexity:} $O(n)$. Each bit is read once and each group
        of 8 bits is processed in constant time.
  \item \textbf{Space complexity:} $O(n)$ to store all bits. This could be reduced to
        $O(1)$ by processing on the fly, accumulating 8 bits at a time before outputting.
\end{itemize}

\end{document}
