\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red},
  numbers=left,
  numberstyle=\tiny\color{gray},
  breaklines=true,
  frame=single,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 2001 -- Double Crypt}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Statement Summary}
A message is encrypted twice using two different simple ciphers, each
parameterized by a key from a known key space:
\[
  \mathrm{ciphertext} = E_{k_2}\bigl(E_{k_1}(\mathrm{plaintext})\bigr).
\]
The key spaces for $k_1$ and $k_2$ each have size up to $M$
(e.g., $M = 2^{24}$). Given a plaintext--ciphertext pair, find
both keys. A brute-force search over all $(k_1, k_2)$ pairs costs
$O(M^2)$, which is prohibitive.

\section{Solution: Meet in the Middle}

\begin{theorem}
The key pair $(k_1, k_2)$ satisfying
$E_{k_2}(E_{k_1}(P)) = C$ can be found in $O(M)$ time and space.
\end{theorem}

\begin{proof}
Rewrite the equation as $E_{k_1}(P) = D_{k_2}(C)$, where $D_{k_2}$
denotes decryption with key $k_2$. Both sides equal a common
\emph{intermediate value} $I$. We enumerate one side, store the
results, and match against the other:

\begin{enumerate}
  \item \textbf{Forward phase.} For every candidate $k_1 \in [0, M)$,
        compute $I = E_{k_1}(P)$ and insert $(I, k_1)$ into a hash table.
  \item \textbf{Backward phase.} For every candidate $k_2 \in [0, M)$,
        compute $I' = D_{k_2}(C)$ and look up $I'$ in the hash table.
        A match yields a candidate pair $(k_1, k_2)$.
\end{enumerate}

Each phase performs $M$ cipher evaluations and $M$ hash-table
operations, giving $O(M)$ total work. The hash table uses $O(M)$ space.
\end{proof}

\paragraph{Handling collisions.}
Multiple $(k_1, k_2)$ pairs may produce the same intermediate value.
Verify each candidate against a second plaintext--ciphertext pair (if
available) to eliminate false positives.

\section{C++ Implementation}

The code below demonstrates the framework using a simple XOR cipher.
Replace \texttt{encrypt}/\texttt{decrypt} with the actual cipher for the
IOI problem; the meet-in-the-middle structure is unchanged.

\begin{lstlisting}
#include <bits/stdc++.h>
using namespace std;

typedef unsigned int uint;

uint encrypt(uint plaintext, uint key) {
    return plaintext ^ key; // replace with actual cipher
}

uint decrypt(uint ciphertext, uint key) {
    return ciphertext ^ key; // replace with actual cipher inverse
}

int main() {
    ios::sync_with_stdio(false);
    cin.tie(nullptr);

    uint plaintext, ciphertext;
    int keyBits;
    cin >> plaintext >> ciphertext >> keyBits;

    uint keySpace = 1u << keyBits;

    // Phase 1: encrypt plaintext with every k1
    unordered_map<uint, uint> table; // intermediate -> k1
    for (uint k1 = 0; k1 < keySpace; k1++) {
        uint mid = encrypt(plaintext, k1);
        table[mid] = k1;
    }

    // Phase 2: decrypt ciphertext with every k2, look up match
    for (uint k2 = 0; k2 < keySpace; k2++) {
        uint mid = decrypt(ciphertext, k2);
        auto it = table.find(mid);
        if (it != table.end()) {
            uint k1 = it->second;
            // Optional: verify against a second pair
            cout << "Key1: " << k1 << "\n";
            cout << "Key2: " << k2 << "\n";
        }
    }

    return 0;
}
\end{lstlisting}

\section{Complexity Analysis}
\begin{itemize}
  \item \textbf{Time:} $O(M)$ where $M = 2^{\texttt{keyBits}}$.
        Each phase performs $M$ cipher evaluations plus $O(1)$-expected-time
        hash-table operations.
  \item \textbf{Space:} $O(M)$ for the hash table.
\end{itemize}

\end{document}
