\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}

\title{ICPC World Finals 2024\\I. Steppe on It}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Problem Summary}

We are given a weighted tree of villages and must place fire engines in exactly $f$ villages. Every village is served by the nearest engine. We want to minimize the maximum distance from any village to its nearest engine.

\section*{Key Observations}

\begin{itemize}[leftmargin=*]
    \item Fix a candidate response time $R$. Then the question becomes: can we place at most $f$ engines so that every village is within distance $R$ of some engine?
    \item This feasibility predicate is monotone. If radius $R$ works, then every larger radius also works. Therefore we can binary search the answer.
    \item For the feasibility check, root the tree arbitrarily. When processing a subtree, only one of two summaries matters:
    \begin{itemize}[leftmargin=*]
        \item the subtree is already fully covered, and we only need the distance from the root of the subtree to the closest engine inside it;
        \item the subtree is not fully covered yet, and we only need the distance to its farthest still-uncovered village.
    \end{itemize}
    \item A greedy postorder DFS is enough: whenever the farthest uncovered village can no longer be covered from the parent, we must place an engine at the current node.
\end{itemize}

\section*{Algorithm}

\begin{enumerate}[leftmargin=*]
    \item Binary search the minimum feasible radius $R$.

    \item For a fixed $R$, run one DFS from the root.

    \item For a node $u$, after processing all children, maintain:
    \begin{itemize}[leftmargin=*]
        \item \texttt{closest\_center}: the minimum distance from $u$ to an engine already placed in a child subtree that is fully covered;
        \item \texttt{farthest\_uncovered}: the maximum distance from $u$ to a village in a child subtree that is still uncovered.
    \end{itemize}

    \item If
    \[
    \texttt{closest\_center} + \texttt{farthest\_uncovered} \le R,
    \]
    then the nearest engine inside the subtree covers every currently uncovered village as well, so the whole subtree becomes fully covered and we return \texttt{closest\_center}.

    \item Otherwise, if $u$ is not the root and
    \[
    \texttt{farthest\_uncovered} > R,
    \]
    then even an engine at the parent would be too far away, so we must place an engine at $u$. We count one engine and return that the subtree is fully covered with nearest-engine distance $0$.

    \item If neither of the two situations above applies, we return that the subtree is still uncovered, with value \texttt{farthest\_uncovered}.

    \item After the DFS, if the root is still uncovered, place one final engine at the root. The check succeeds iff the total number of engines used is at most $f$.
\end{enumerate}

\section*{Correctness Proof}

We prove that the algorithm returns the correct answer.

\paragraph{Lemma 1.}
During the feasibility check for radius $R$, if a node $u$ returns the state ``uncovered with value $d$'', then every uncovered village in the subtree of $u$ is at distance at most $d$ from $u$, and at least one such village is at distance exactly $d$.

\paragraph{Proof.}
This follows directly from the DFS definition. Whenever a child returns uncovered, we propagate the maximum uncovered distance upward by adding the edge length. Thus the stored value is exactly the farthest uncovered distance from the current node. \qed

\paragraph{Lemma 2.}
If at a non-root node $u$ we have \texttt{farthest\_uncovered} $> R$, then every feasible solution for radius $R$ must place an engine at $u$ or inside the subtree of $u$.

\paragraph{Proof.}
By Lemma 1 there is an uncovered village in the subtree of $u$ at distance more than $R$ from $u$. Any engine outside the subtree of $u$ must be even farther away, because every path from that village to the outside passes through $u$. Hence no engine above $u$ can cover that village. Therefore some engine must be placed at $u$ or below it. The greedy choice of placing it exactly at $u$ is optimal because it is the highest possible placement that still covers the problematic part of the subtree. \qed

\paragraph{Lemma 3.}
If \texttt{closest\_center} $+$ \texttt{farthest\_uncovered} $\le R$ at node $u$, then all uncovered villages in the subtree of $u$ are in fact already covered by the nearest engine in that subtree.

\paragraph{Proof.}
Every uncovered village is at distance at most \texttt{farthest\_uncovered} from $u$, while the nearest engine is at distance \texttt{closest\_center} from $u$. The unique path between them goes through $u$, so their mutual distance is at most the sum of the two values, which is at most $R$. Hence all those villages are actually covered, and the whole subtree is fully covered. \qed

\paragraph{Theorem.}
The algorithm outputs the correct answer for every valid input.

\paragraph{Proof.}
For a fixed radius $R$, Lemma 2 shows that whenever the DFS places an engine, every feasible solution must also place an engine somewhere in that subtree, so the greedy action is necessary. Lemma 3 shows that whenever the DFS declares a subtree fully covered, this is correct. Therefore the DFS computes the minimum number of engines needed for radius $R$.

Since the feasibility predicate is monotone in $R$, binary search returns the smallest radius for which the DFS uses at most $f$ engines. This is exactly the minimum guaranteed response time. \qed

\section*{Complexity Analysis}

Each feasibility check is one DFS, so it costs $O(n)$. The answer is found by binary search over an integer range whose upper bound is the total tree length, so the total running time is $O(n \log W)$, where $W$ is that upper bound. The memory usage is $O(n)$.

\section*{Implementation Notes}

\begin{itemize}[leftmargin=*]
    \item Edge lengths and path lengths require \texttt{long long}.
    \item The DFS is recursive in the write-up, but any equivalent postorder implementation is fine. The code roots the tree at village $1$.
    \item If the root is still uncovered after processing all children, we must count one additional engine at the root.
\end{itemize}

\end{document}
