\section*{Overview}

KMP is the classical linear-time pattern matcher built on the prefix function. Its main lesson is not ``use this one
trick''. Its main lesson is that after a mismatch, we should not restart from scratch if some of the matched prefix is
still usable.

That same prefix-function viewpoint also explains borders, periodicity, and several other string facts that show up in
contest problems.

\section*{When to Use It}

Use it when:

\begin{itemize}
  \item you need all occurrences of one pattern in one text,
  \item you want a deterministic linear-time matcher,
  \item hashing would be unnecessary or too fragile,
  \item border information or periodicity is relevant, even if full matching is not.
\end{itemize}

For one casual substring check, the standard library is often enough. KMP matters when the prefix-function structure
itself is useful.

\section*{Core Idea}

For each position \(i\), the prefix function \(\pi[i]\) is the length of the longest proper prefix of the string that
is also a suffix ending at \(i\).

That ``prefix equals suffix'' object is called a \textbf{border}. KMP uses borders as fallback positions: after a
mismatch, the only remaining candidates are borders of what has already matched.

\section*{Key Insight}

The invariant during the text scan is:
\textbf{the current matched length is the longest prefix of the pattern that is also a suffix of the processed text}.

So if a mismatch happens after matching \(k\) characters, the next sensible candidate is not ``start over''. It is the
longest border of that matched prefix, whose length is \(\pi[k - 1]\).

That is why no text character needs to be revisited from scratch.

\section*{Operations / Main Technique}

The standard workflow is:

\begin{itemize}
  \item compute the prefix function of the pattern,
  \item scan the text left to right,
  \item on mismatch, fall back through prefix-function links,
  \item on full match, report the occurrence and continue from the next border.
\end{itemize}

\paragraph{Worked example.}
If the pattern prefix \(\texttt{abab}\) has matched and the next text character mismatches, KMP does not forget
everything. The matched prefix has border \(\texttt{ab}\), so matching can continue as if those two characters had
already been confirmed.

\section*{Correctness Intuition}

At every step, the matched length is the longest possible viable prefix length. If a mismatch occurs, any longer
candidate is impossible, so the next candidate must be the longest border of the current matched prefix. Repeating that
step eventually finds the longest remaining candidate that could still continue the match.

Because the matched length only moves forward or falls back along prefix-function links, the total amount of fallback
work is linear.

\section*{Complexity Analysis}

\begin{itemize}
  \item prefix function of a length-\(n\) pattern: \(O(n)\),
  \item scan of a length-\(m\) text: \(O(m)\),
  \item total matching time: \(O(n + m)\),
  \item memory: \(O(n)\).
\end{itemize}

\section*{Implementation}

The reference code separates:

\begin{itemize}
  \item \texttt{prefix\_function(pattern)},
  \item \texttt{kmp\_search(text, pattern)}.
\end{itemize}

That split matters because many problems only need the prefix function itself. Matching is just one of its most famous
applications.

\section*{Common Pitfalls}

\begin{itemize}
  \item Confusing the prefix function with the Z-function.
  \item Forgetting that \(\pi[i]\) is a \emph{proper} prefix length.
  \item Mishandling the empty-pattern edge case.
  \item Writing the fallback loop incorrectly and skipping valid borders.
  \item Treating KMP as only a matching tool and missing border/periodicity applications.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item prefix-function automaton style transitions,
  \item border enumeration,
  \item periodicity tests,
  \item counting occurrences of every prefix,
  \item comparison with the Z-function, which stores a different but related prefix-match view.
\end{itemize}

The Z-function often feels cleaner for some pattern tasks, but KMP is still the more natural language for borders and
periodicity.

\section*{Practice Problems}

\begin{itemize}
  \item Find all pattern occurrences in a long text.
  \item Compute all borders of a string.
  \item Detect the smallest repeating period of a string.
  \item Problems that combine a small automaton with digit DP or DP on strings.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/string/prefix-function.html}{cp-algorithms: Prefix function and KMP}.
  \item \href{https://atcoder.github.io/ac-library/production/document_en/string.html}{AtCoder Library string documentation}.
  \item \href{https://usaco.guide/CPH.pdf}{Competitive Programmer's Handbook}.
\end{itemize}
