\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}

\title{ICPC World Finals 2025\\K. Treasure Map}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Problem Summary}

We know some depths on an $n \times m$ grid and must extend them to a full map such that on every unit
square the two possible diagonal triangulations produce the same linear interpolation. Among all valid
extensions with nonnegative depths everywhere, we want the smallest possible depth at the treasure
point.

\section*{Key Observations}

\begin{itemize}[leftmargin=*]
    \item On one unit square with corners $A,B,C,D$ in cyclic order, the two diagonal interpolations agree
    everywhere if and only if
    \[
    d(A)+d(C)=d(B)+d(D).
    \]
    \item This is exactly the condition that the discrete mixed second difference is zero. Therefore every
    valid map has the form
    \[
    d(x,y)=R_x + C_y
    \]
    for some row potentials $R_x$ and column potentials $C_y$.
    \item Every known depth gives one equation $R_x + C_y = d$.
    This is a bipartite graph on the row vertices and column vertices.
    Inside one connected component, all potentials are determined up to one free shift
    \[
    R \leftarrow R+t,\qquad C \leftarrow C-t.
    \]
    \item Nonnegativity means $R_x + C_y \ge 0$ for every grid point.
    For one connected component containing both row and column vertices, the quantity
    \[
    \min R_x + \min C_y
    \]
    is invariant under the shift above; if it is negative, the instance is impossible.
    \item To minimize the treasure depth, if the treasure row and treasure column are in different connected
    components, we can shift those two components independently as far downward as nonnegativity
    allows.
\end{itemize}

\section*{Algorithm}

\begin{enumerate}[leftmargin=*]
    \item Build a bipartite graph with row vertices $1,\dots,n$ and column vertices $n+1,\dots,n+m$.
    For every known point $(x,y,d)$, add edges encoding $R_x + C_y = d$.
    \item Run BFS/DFS on each connected component.
    Fix one starting potential to $0$ and propagate all other potentials using the equations.
    If we ever derive two different values for the same vertex, output \texttt{impossible}.
    \item For every connected component, record:
    the minimum row potential inside it and the minimum column potential inside it.
    If the component contains both sides and their sum is negative, output \texttt{impossible}.
    \item Let the treasure row be $R_{t_x}$ and the treasure column be $C_{t_y}$.
    If they lie in the same component, their sum is fixed.
    Otherwise, shift the row component down to make its minimum row potential $0$, and shift the column
    component down to make its minimum column potential $0$.
    The resulting treasure depth is the minimum possible one.
\end{enumerate}

\section*{Correctness Proof}

We prove that the algorithm returns the correct answer.

\paragraph{Lemma 1.}
The interpolation condition on every unit square is equivalent to the existence of row and column
potentials with $d(x,y)=R_x+C_y$.

\paragraph{Proof.}
If $d(x,y)=R_x+C_y$, then for any square with corners $(x,y),(x+1,y),(x+1,y+1),(x,y+1)$ we have
\[
d(x,y)+d(x+1,y+1)=R_x+C_y+R_{x+1}+C_{y+1}
                  =d(x+1,y)+d(x,y+1),
\]
so the diagonal-interpolation condition holds.

Conversely, if the mixed second difference is zero on every unit square, fix $R_1=0$ and define
$C_y=d(1,y)$ and $R_x=d(x,1)-C_1$.
The zero second-difference condition then propagates inductively to all grid points and yields
$d(x,y)=R_x+C_y$. \qed

\paragraph{Lemma 2.}
Within one connected component of the known-point graph, all potentials are determined up to the shift
$R \leftarrow R+t$, $C \leftarrow C-t$.

\paragraph{Proof.}
Every known point gives one equation of the form $R_x + C_y = d$.
Along any path in the bipartite graph, once one endpoint value is fixed, all others are forced.
Changing all row values by $+t$ and all column values by $-t$ preserves every such equation, and this is
the only remaining freedom in a connected component. \qed

\paragraph{Lemma 3.}
For one connected component, nonnegativity is feasible if and only if
\[
\min R_x + \min C_y \ge 0.
\]

\paragraph{Proof.}
Under the shift from Lemma 2, every sum $R_x+C_y$ inside the same component is unchanged.
Hence the smallest possible value anywhere in the component is exactly
$\min R_x + \min C_y$, which is therefore shift-invariant.
If this number is negative, some grid point in the component is necessarily negative in every extension.
If it is nonnegative, then all sums in the component are nonnegative already. \qed

\paragraph{Theorem.}
The algorithm outputs the correct answer for every valid input.

\paragraph{Proof.}
By Lemma 1, every valid map is exactly a choice of row and column potentials satisfying the known-point
equations. Lemma 2 describes all freedom left after propagating those equations, and Lemma 3 gives the
exact feasibility test for nonnegativity.

If the treasure row and treasure column are in the same connected component, their sum is invariant under
all allowed shifts, so the treasure depth is fixed.
If they are in different components, the two components can be shifted independently, and pushing each
one downward until one of its minima becomes $0$ clearly minimizes the treasure sum.
Therefore the value printed by the algorithm is exactly the smallest possible treasure depth. \qed

\section*{Complexity Analysis}

The graph has $n+m$ vertices and $k$ edges. The BFS/DFS over all components runs in $O(n+m+k)$
time and uses $O(n+m+k)$ memory.

\section*{Implementation Notes}

\begin{itemize}[leftmargin=*]
    \item The implementation stores row vertices as $1,\dots,n$ and column vertices as $n+1,\dots,n+m$.
    \item Potentials are kept as 64-bit integers; the statement guarantees that the optimum is integral.
\end{itemize}

\end{document}
