\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

\title{Problem 242: Odd Triplets}
\author{Project Euler Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

Define $f(N)$ as the number of triplets $(a,b,c)$ with $0<a\le b\le c$, $a+b+c\le N$, and $a\oplus b\oplus c=0$. Find $f(10^{12})\bmod 10^9$.

\section{Mathematical Development}

\begin{definition}
The set of valid XOR-zero bit patterns is $\mathcal{V}=\{(0,0,0),(0,1,1),(1,0,1),(1,1,0)\}$.
\end{definition}

\begin{theorem}[XOR-zero bit classification]\label{thm:xor}
$a\oplus b\oplus c=0$ if and only if at every bit position the triple of bits belongs to~$\mathcal{V}$.
\end{theorem}

\begin{proof}
XOR operates independently at each bit. The XOR of three bits is zero iff an even number (0 or~2) are~1, which yields exactly the four patterns in~$\mathcal{V}$.
\end{proof}

\begin{lemma}[Bit-sum structure]\label{lem:bitsum}
If $a\oplus b\oplus c=0$, then at each bit position the arithmetic sum $s_i\in\{0,2\}$. Consequently $a+b+c=\sum_i s_i\cdot 2^i$ is always even.
\end{lemma}

\begin{proof}
Each pattern in $\mathcal{V}$ sums to either~0 or~2. Since every term $s_i\cdot 2^i$ is divisible by~2, so is the total.
\end{proof}

\begin{theorem}[Digit DP formulation]\label{thm:dp}
Write $N=\sum_{i=0}^{L}n_i\cdot 2^i$ in binary. Process bits from position~$L$ down to~$0$, maintaining state $(\mathrm{carry},\mathrm{tight},\mathrm{ord}_{ab},\mathrm{ord}_{bc})$ where:
\begin{itemize}
\item $\mathrm{carry}\in\{0,1\}$: carry from the addition $a+b+c$;
\item $\mathrm{tight}\in\{0,1\}$: whether $a+b+c$ remains bounded by~$N$;
\item $\mathrm{ord}_{ab},\mathrm{ord}_{bc}\in\{=,<\}$: ordering status for $a\le b$ and $b\le c$.
\end{itemize}
At each bit, enumerate all patterns in $\mathcal{V}$, updating carry, tightness, and ordering. Accept at $i=-1$ iff $\mathrm{carry}=0$.
\end{theorem}

\begin{proof}
\textbf{(i)~Exhaustiveness.} All four valid XOR-zero patterns are tried at each bit.

\textbf{(ii)~Ordering correctness.} If $\mathrm{ord}_{ab}={=}$, enforce $a_i\le b_i$; transition to ${<}$ when $a_i<b_i$. Once ${<}$, any $(a_i,b_i)$ is valid since $a<b$ is established at higher bits. Analogously for $\mathrm{ord}_{bc}$.

\textbf{(iii)~Sum constraint.} Carries propagate correctly; tightness ensures $a+b+c\le N$.

\textbf{(iv)~Termination.} $\mathrm{carry}=0$ at position $-1$ confirms no overflow.
\end{proof}

\section{Algorithm}

\begin{verbatim}
dp = {(0, 1, EQUAL, EQUAL): 1}
for pos = L downto 0:
    for each state and each pattern in V:
        check ordering, compute carry/digit, update tightness
        accumulate into next-level dp
result = sum(counts with carry=0) - 1   # subtract (0,0,0)
\end{verbatim}

\section{Complexity Analysis}

\begin{itemize}
\item \textbf{Time:} $O(L\cdot|S|\cdot|\mathcal{V}|)=O(40\times 16\times 4)=O(2560)$.
\item \textbf{Space:} $O(|S|)=O(16)=O(1)$.
\end{itemize}

\section{Answer}

\[
\boxed{997104142249036713}
\]

\end{document}
