\section*{Overview}

Suffix array is the lightest full suffix structure I regularly want in contests. It orders all suffixes of a string
lexicographically, which makes substring comparison, lexicographic queries, and several repeated-pattern tasks much more
manageable.

\section*{When to Use It}

Use a suffix array when:

\begin{itemize}
  \item lexicographic order of suffixes matters,
  \item many substring comparisons or pattern queries are needed,
  \item the problem wants structure stronger than hashing but simpler than a suffix tree.
\end{itemize}

\section*{Core Idea}

The suffix array \(\texttt{sa}\) stores the starting indices of the suffixes in sorted order. The standard doubling
construction sorts suffixes by the first \(2^k\) characters, doubling that compared prefix length each round.

\section*{Key Insight}

Two suffixes can be compared by pairs of class ids from the previous round. That turns the expensive ``compare long
strings'' step into sorting small tuples that summarize larger and larger prefixes.

\section*{Operations / Main Technique}

\begin{itemize}
  \item build \(\texttt{sa}\) by doubling,
  \item build the LCP array with Kasai's algorithm,
  \item use binary search on the suffix array for pattern lookup,
  \item reduce repeated-substring tasks to neighbors in suffix order.
\end{itemize}

\paragraph{Worked example.}
If two suffixes share a long common prefix, they appear near each other in suffix-array order. That is why repeated
substring questions usually look at adjacent suffixes and the LCP array rather than arbitrary pairs.

\section*{Correctness Intuition}

After round \(k\), the class ids correctly describe the order of all length-\(2^k\) prefixes. Sorting the pair
\((class[i], class[i + 2^k])\) is therefore exactly sorting the first \(2^{k+1}\) characters. Once the compared length
exceeds the string length, the order is the full suffix order.

\section*{Complexity Analysis}

The doubling implementation is \(O(n \log n)\), and Kasai's LCP construction is \(O(n)\).

\section*{Implementation}

The code returns both:

\begin{itemize}
  \item \texttt{sa},
  \item \texttt{lcp}.
\end{itemize}

That pair is enough for most contest uses of the suffix array.

\section*{Common Pitfalls}

\begin{itemize}
  \item Forgetting to append a unique terminator smaller than all real characters.
  \item Using a suffix array when a simpler KMP, Z-function, or hash-based solution already works.
  \item Misaligning the LCP array: \(\texttt{lcp[i]}\) belongs to adjacent suffixes in suffix-array order, not original order.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Sparse table or RMQ over LCP for fast LCP queries between arbitrary suffixes.
  \item Suffix automaton when online extension or substring counting is more natural.
  \item SA-IS or linear-time construction, which is powerful but often unnecessary in contest code.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Count distinct substrings.
  \item Find the lexicographically smallest rotation.
  \item Pattern lookup and repeated-substring tasks using SA and LCP.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/string/suffix-array.html}{cp-algorithms: Suffix Array}.
  \item \href{https://usaco.guide/adv/suffix-array?lang=cpp}{USACO Guide: Suffix Array}.
  \item \href{https://codeforces.com/catalog?locale=en}{Codeforces Catalog}.
\end{itemize}
