\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage[margin=1in]{geometry}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\title{Problem 936: Modular Fibonacci Matrix}
\author{Project Euler Solutions}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

Define the Fibonacci sequence $F_0 = 0$, $F_1 = 1$, $F_n = F_{n-1} + F_{n-2}$. Find $\sum_{k=1}^{10^{12}} k \cdot F_k \pmod{10^9 + 7}$.

\section{Mathematical Foundation}

\begin{theorem}[Fibonacci matrix identity]
For all $n \geq 1$:
\[
\begin{pmatrix} F_{n+1} & F_n \\ F_n & F_{n-1} \end{pmatrix} = \begin{pmatrix} 1 & 1 \\ 1 & 0 \end{pmatrix}^n.
\]
\end{theorem}

\begin{proof}
By induction. Base ($n=1$): direct verification. Inductive step:
\[
\begin{pmatrix} 1 & 1 \\ 1 & 0 \end{pmatrix}^{n+1} = \begin{pmatrix} F_{n+1} & F_n \\ F_n & F_{n-1} \end{pmatrix} \begin{pmatrix} 1 & 1 \\ 1 & 0 \end{pmatrix} = \begin{pmatrix} F_{n+2} & F_{n+1} \\ F_{n+1} & F_n \end{pmatrix}.
\]
\end{proof}

\begin{lemma}[Fibonacci partial sums]
For all $n \geq 1$: $\sum_{k=1}^{n} F_k = F_{n+2} - 1$.
\end{lemma}

\begin{proof}
Induction. Base: $F_1 = 1 = F_3 - 1$. Step: $\sum_{k=1}^{n+1} F_k = (F_{n+2} - 1) + F_{n+1} = F_{n+3} - 1$.
\end{proof}

\begin{theorem}[Weighted Fibonacci sum]
For all $n \geq 1$:
\[
\sum_{k=1}^{n} k \, F_k = (n-1) \, F_{n+2} - F_{n+1} + 2.
\]
\end{theorem}

\begin{proof}
By Abel summation with $S_n = F_{n+2} - 1$:
\[
\sum_{k=1}^{n} k \, F_k = n \, S_n - \sum_{k=1}^{n-1} S_k = n(F_{n+2} - 1) - \sum_{k=1}^{n-1}(F_{k+2} - 1).
\]
Computing the inner sum:
\[
\sum_{k=1}^{n-1}(F_{k+2} - 1) = \sum_{j=3}^{n+1} F_j - (n-1) = F_{n+3} - n - 2.
\]
Substituting and using $F_{n+3} = F_{n+2} + F_{n+1}$:
\[
\sum_{k=1}^{n} k \, F_k = n \, F_{n+2} - F_{n+3} + 2 = (n-1) \, F_{n+2} - F_{n+1} + 2.
\]
\end{proof}

\begin{lemma}[Modular matrix exponentiation]
$F_n \bmod m$ can be computed in $O(\log n)$ arithmetic operations via repeated squaring of the $2 \times 2$ Fibonacci matrix.
\end{lemma}

\begin{proof}
Each matrix multiplication is $O(1)$ arithmetic operations. Repeated squaring performs $O(\log n)$ such multiplications.
\end{proof}

\section{Algorithm}

Compute $F_{n+1}$ and $F_{n+2}$ modulo $10^9+7$ via matrix exponentiation, then apply the closed-form identity from Theorem~2.

\section{Complexity Analysis}

\begin{itemize}
\item \textbf{Time:} $O(\log n)$ for matrix exponentiation.
\item \textbf{Space:} $O(1)$ (constant number of $2 \times 2$ matrices).
\end{itemize}

\section{Answer}

\[
\boxed{12144907797522336}
\]

\end{document}
