\section*{Overview}

Point-in-polygon is the routine that turns polygon input into actual queries. In contests the goal is usually modest:
given a simple polygon and many query points, classify each point as outside, on the boundary, or strictly inside.

\section*{When to Use It}

Use this when:

\begin{itemize}
  \item the input polygon is fixed and many query points follow,
  \item the statement distinguishes inside from boundary,
  \item you want a simple, reliable predicate rather than heavy geometry machinery.
\end{itemize}

\section*{Core Idea}

The standard parity test casts a ray from the query point and counts how many polygon edges cross it. If the number of
crossings is odd, the point is inside; if even, it is outside.

Before applying parity, check whether the point lies on an edge. That boundary case should be handled explicitly.

\section*{Key Insight}

The difficult part is not the odd/even rule. It is choosing a crossing convention that does not double-count vertices.
The common fix is to count an edge only when its endpoints straddle the horizontal line in a half-open way.

\section*{Operations / Main Technique}

\begin{itemize}
  \item test every edge for boundary contact,
  \item apply a ray-crossing parity count,
  \item return a three-way classification.
\end{itemize}

\paragraph{Worked example.}
If a horizontal ray from the query point crosses the polygon boundary three times, the point is inside. If it crosses
zero or two times, it is outside. If the point lies on one of the edges, the parity rule is skipped.

\section*{Correctness Intuition}

A ray from an outside point enters and leaves the polygon in pairs, so the crossing count is even. A ray from an
inside point has one unmatched exit, so the count is odd. The boundary check removes degenerate cases where the point
lies exactly on the polygon.

\section*{Complexity Analysis}

For a polygon with \(n\) vertices, one query takes \(O(n)\) time and \(O(1)\) extra memory.

\section*{Implementation}

The sample code returns:

\begin{itemize}
  \item \texttt{0} for outside,
  \item \texttt{1} for boundary,
  \item \texttt{2} for inside.
\end{itemize}

For many queries on the same polygon, that linear routine is often enough unless the constraints are very large.

\section*{Common Pitfalls}

\begin{itemize}
  \item Double-counting vertices in the ray-crossing loop.
  \item Forgetting the boundary case and classifying it by parity.
  \item Assuming the polygon is convex when the problem does not say so.
  \item Using floating-point comparisons when an integer boundary test is available.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Winding-number interpretation instead of parity.
  \item Faster query structures for many points against one polygon.
  \item Convex-polygon queries in \(O(\log n)\) after preprocessing.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Classify query points relative to a simple polygon.
  \item Count how many points lie strictly inside.
  \item Combine point-in-polygon with polygon area or intersection tasks.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/geometry/point-in-convex-polygon.html}{cp-algorithms: Point in Polygon related geometry routines}.
  \item \href{https://usaco.guide/plat/geo-prims}{USACO Guide: Geometry Primitives}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
