\section*{Overview}

The treap is the balanced binary search tree I find easiest to turn into contest code. It combines two invariants:
the keys satisfy the binary-search-tree order, while random priorities satisfy the heap order.

That combination gives a tree that behaves like a random BST. In practice, it means I get split and merge almost for
free, which is why treaps show up in ordered sets, implicit sequences, and many ``cut this interval, move that
interval'' problems.

\section*{When to Use It}

Use it when:

\begin{itemize}
  \item you need an ordered set with insert and erase,
  \item split and merge are the cleanest primitive operations,
  \item an implicit sequence representation would be useful later,
  \item you want randomized balancing instead of amortized balancing.
\end{itemize}

If the task only needs an ordinary ordered set, the STL may already be enough. Treaps become attractive when you need
direct control of the tree structure itself.

\section*{Core Idea}

Each node stores:

\begin{itemize}
  \item a \textbf{key} for BST order,
  \item a random \textbf{priority} for heap order.
\end{itemize}

So the inorder traversal is sorted by key, while priorities decrease from parent to child.

The useful consequence is that once priorities are fixed, the treap shape is determined. Random priorities make that
shape behave like a randomly built BST, which keeps the expected height logarithmic.

\section*{Key Insight}

The treap is easiest to reason about through \textbf{split} and \textbf{merge}, not through raw rotations.

\begin{itemize}
  \item \textbf{split(root, key)} returns two treaps: keys \(< key\) and keys \(\ge key\),
  \item \textbf{merge(left, right)} assumes all keys in the left treap are smaller and rebuilds one valid treap.
\end{itemize}

Once those two routines are correct, insert and erase are usually only a few lines of logic around them.

\section*{Operations / Main Technique}

The standard ordered-set version supports:

\begin{itemize}
  \item \textbf{split},
  \item \textbf{merge},
  \item \textbf{insert},
  \item \textbf{erase},
  \item \textbf{kth} or order-statistic queries if subtree sizes are stored.
\end{itemize}

\paragraph{Worked example.}
To insert \(x\), split by \(x\), create one node with key \(x\), then merge left + node + right. To erase \(x\), find
the node, remove it, and merge its left and right children back together.

That is why treaps feel modular: the balancing is hidden inside merge rather than spread throughout many case splits.

\section*{Correctness Intuition}

BST order is preserved because split partitions keys correctly and merge is only called when every key in the left
treap is smaller than every key in the right treap.

Heap order is preserved because merge always picks the root with higher priority, then recursively merges the remaining
subtree on the correct side. Since both subtrees were already valid treaps, the combined result is a valid treap too.

\section*{Complexity Analysis}

With random priorities:

\begin{itemize}
  \item split, merge, insert, erase, kth: \(O(\log n)\) expected,
  \item memory: \(O(n)\).
\end{itemize}

The worst case still exists, but randomized priorities make it unlikely enough for normal contest use.

\section*{Implementation}

The reference code keeps:

\begin{itemize}
  \item explicit node pointers,
  \item random priorities,
  \item subtree sizes,
  \item split / merge,
  \item insert / erase / kth.
\end{itemize}

That is the version I like as a base because it scales naturally to implicit treaps, where indices replace keys and
interval operations become split/merge chains.

\section*{Common Pitfalls}

\begin{itemize}
  \item Forgetting to pull subtree size after recursive changes.
  \item Merging trees whose key ranges overlap.
  \item Using a weak random source and then trusting the shape too much.
  \item Mixing ``duplicates allowed'' and ``unique keys'' rules halfway through the implementation.
  \item Writing implicit-treap code before the plain ordered-set version feels stable.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item implicit treap for sequence manipulation,
  \item lazy tags for range reverse, range add, or range assign,
  \item treap with duplicate counts,
  \item persistent treap variants,
  \item union of two treaps in \(O(m \log(n/m))\) expected time.
\end{itemize}

Compared with splay trees, treaps are randomized rather than amortized. I usually reach for treap first when I want
something predictable to code quickly, and for splay when I specifically want access-based restructuring or link-cut
trees later.

\section*{Practice Problems}

\begin{itemize}
  \item Ordered set with insert, erase, predecessor, successor, and k-th queries.
  \item Cut-and-paste sequence problems with an implicit treap.
  \item Interval reverse or rotate operations after adding lazy tags.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/data_structures/treap.html}{cp-algorithms: Treap}.
  \item \href{https://usaco.guide/adv/treaps}{USACO Guide: Treaps}.
  \item \href{https://github.com/kth-competitive-programming/kactl}{KACTL}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
