\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{listings}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{margin=2.5cm}

\newtheorem{lemma}{Lemma}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\bfseries,
  breaklines=true,
  numbers=left,
  numberstyle=\tiny,
  frame=single,
  tabsize=2
}

\title{IOI 2019 -- Arranging Shoes}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Summary}
There are $2n$ shoes: $n$ left shoes (represented by $-i$) and $n$ right shoes ($+i$) for sizes $i = 1, \ldots, n$. Rearrange using the minimum number of adjacent swaps so that each pair is adjacent with the left shoe on the left.

\section{Solution}

\subsection{Greedy Strategy}
Process positions left to right. For each even position $i$ (start of a pair slot):
\begin{enumerate}
  \item Let shoe $s$ be at position $i$. If $s > 0$ (right shoe), its matching left shoe $-s$ is somewhere to the right; conceptually swap them (costs 1 extra swap) so the left shoe is at position $i$.
  \item Find the matching shoe and bring it to position $i+1$ by counting the number of \emph{active} (not yet matched) positions between them.
  \item Mark both positions as used.
\end{enumerate}

\subsection{Counting with a Fenwick Tree}
A Fenwick tree tracks active positions. The cost to bring the match from position $j$ to the slot adjacent to position $i$ is the number of active positions strictly between $i$ and $j$.

\begin{lemma}
This greedy produces the minimum number of swaps.
\end{lemma}
\begin{proof}[Proof sketch]
At each step, the pair occupying the leftmost unfilled slot is fixed. The greedy matches each pair optimally: bringing the partner with the minimum number of intermediate active elements. Since completed pairs never need to be disturbed again, this is optimal.
\end{proof}

\section{Implementation}

\begin{lstlisting}
#include <bits/stdc++.h>
using namespace std;

struct BIT {
    int n;
    vector<int> tree;
    BIT(int n) : n(n), tree(n + 1, 0) {}
    void update(int i, int v) { for (i++; i <= n; i += i & -i) tree[i] += v; }
    int query(int i) { int s = 0; for (i++; i > 0; i -= i & -i) s += tree[i]; return s; }
    int query(int l, int r) { return query(r) - (l > 0 ? query(l - 1) : 0); }
};

long long count_swaps(vector<int> S) {
    int n2 = S.size(), n = n2 / 2;
    map<int, queue<int>> pos;  // shoe value -> queue of positions
    for (int i = 0; i < n2; i++) pos[S[i]].push(i);

    BIT bit(n2);
    for (int i = 0; i < n2; i++) bit.update(i, 1);

    vector<bool> used(n2, false);
    long long ans = 0;

    for (int i = 0; i < n2; i++) {
        if (used[i]) continue;

        int shoe = S[i];
        int matchPos;

        if (shoe < 0) {
            // Left shoe: find matching right shoe
            pos[-shoe].pop();  // consume position i from right queue? No, consume from left.
            pos[shoe].pop();   // consume position i from left queue
            matchPos = pos[-shoe].front(); pos[-shoe].pop();
        } else {
            // Right shoe at left slot: extra swap needed
            pos[shoe].pop();   // consume position i
            matchPos = pos[-shoe].front(); pos[-shoe].pop();
            ans++;  // one swap to put left before right
        }

        // Cost = number of active positions strictly between i and matchPos
        int lo = min(i, matchPos), hi = max(i, matchPos);
        ans += bit.query(lo + 1, hi - 1);

        used[i] = true; used[matchPos] = true;
        bit.update(i, -1); bit.update(matchPos, -1);
    }

    return ans;
}
\end{lstlisting}

\section{Complexity Analysis}
\begin{itemize}
  \item \textbf{Time}: $O(n \log n)$ -- each shoe processed once with $O(\log n)$ Fenwick tree operations.
  \item \textbf{Space}: $O(n)$.
\end{itemize}

\end{document}
