\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}

\title{ICPC World Finals 2021\\E. Hand of the Free Marked}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Problem Summary}

From the deck, $k$ cards are chosen uniformly at random. The assistant hides one of them and lays out the
other $k-1$ cards in some order. The magician sees:
\begin{itemize}[leftmargin=*]
    \item the exact identities of the $k-1$ visible cards;
    \item the back marking of the hidden card.
\end{itemize}

We want the highest possible probability that the magician can identify the hidden card.

\section*{Split by Mark Pattern}

Let the mark counts be $c_1, c_2, \dots, c_m$. For one random choice of $k$ cards, only the multiset of
marks matters. Write that multiset as
\[
    x_1, x_2, \dots, x_m,
\]
where $x_i$ cards of mark $i$ were selected and
\[
    x_1 + x_2 + \cdots + x_m = k.
\]

Once the cards are on the table, the magician knows this vector exactly: the visible cards reveal their
marks, and the hidden card reveals its back mark. Therefore different vectors $(x_1,\dots,x_m)$ are
completely independent subproblems, and the assistant and magician may use a different optimal strategy
for each one.

\section*{Bipartite Graph for One Pattern}

Fix one pattern $x$.

\paragraph{Right side.}
Each vertex is an actual $k$-card subset having this mark pattern. The number of such subsets is
\[
    R(x) = \prod_{i=1}^m \binom{c_i}{x_i}.
\]

\paragraph{Left side.}
Each vertex is one possible table state:
\begin{itemize}[leftmargin=*]
    \item which mark the hidden card has;
    \item which $k-1$ cards are visible;
    \item in what order they are shown.
\end{itemize}

If the hidden card has mark $h$, then the visible cards contain
\[
    x_h - 1
\]
cards of mark $h$ and $x_i$ cards of every other mark $i$. So the number of left vertices with hidden
mark $h$ is
\[
    (k-1)! \binom{c_h}{x_h - 1} \prod_{i \ne h} \binom{c_i}{x_i}.
\]
Summing over all possible hidden marks gives
\[
    L(x) = (k-1)! \sum_{h : x_h > 0}
    \binom{c_h}{x_h - 1} \prod_{i \ne h} \binom{c_i}{x_i}.
\]

\paragraph{Edges.}
A right vertex is adjacent to exactly the table states that can be produced from that selected $k$-set.
Choosing a strategy is exactly choosing, for as many right vertices as possible, distinct adjacent left
vertices. That is a matching problem.

By the same symmetry argument used in the classical Fitch Cheney trick, this graph always has a matching
that saturates the smaller side. Therefore, for this pattern, the number of successful selected sets is
\[
    \min(L(x), R(x)).
\]

\section*{A Useful Simplification}

Divide $L(x)$ by $R(x)$:
\[
    \frac{L(x)}{R(x)}
    = (k-1)! \sum_{h : x_h > 0}
    \frac{\binom{c_h}{x_h - 1}}{\binom{c_h}{x_h}}
    = (k-1)! \sum_{h : x_h > 0}
    \frac{x_h}{c_h - x_h + 1}.
\]

So the success probability conditioned on pattern $x$ is
\[
    \min\!\left(1,\,
    (k-1)! \sum_{h : x_h > 0}
    \frac{x_h}{c_h - x_h + 1}\right).
\]

The probability of obtaining pattern $x$ in the random draw is the multivariate hypergeometric value
\[
    \frac{R(x)}{\binom{n}{k}}.
\]

Therefore the final answer is
\[
    \sum_{x_1+\cdots+x_m=k}
    \frac{R(x)}{\binom{n}{k}}
    \cdot
    \min\!\left(1,\,
    (k-1)! \sum_{h : x_h > 0}
    \frac{x_h}{c_h - x_h + 1}\right),
\]
with the additional restriction $0 \le x_i \le c_i$.

\section*{Algorithm}

\begin{enumerate}[leftmargin=*]
    \item Enumerate all vectors $(x_1,\dots,x_m)$ such that
    \[
        x_1+\cdots+x_m=k, \qquad 0 \le x_i \le c_i.
    \]
    Since $k,m \le 10$, this search space is tiny.
    \item For each vector:
    \begin{itemize}[leftmargin=*]
        \item compute
        \[
            R(x) = \prod_i \binom{c_i}{x_i};
        \]
        \item compute the conditional success factor
        \[
            \min\!\left(1,\,
            (k-1)! \sum_h \frac{x_h}{c_h - x_h + 1}\right);
        \]
        \item add the corresponding weighted contribution to the answer.
    \end{itemize}
\end{enumerate}

All binomial coefficients only need $r \le 10$, so they can be evaluated directly in $O(r)$ time using
long doubles.

\section*{Correctness Proof}

We prove that the algorithm returns the optimal success probability.

\paragraph{Lemma 1.}
Different mark patterns $(x_1,\dots,x_m)$ can be optimized independently.

\paragraph{Proof.}
The magician sees all visible cards and also the mark on the hidden card. Hence the full mark multiset of
the chosen $k$ cards is known before any final guess is made. So the assistant and magician may agree on
a separate strategy for each mark pattern without any ambiguity. \qed

\paragraph{Lemma 2.}
For a fixed mark pattern $x$, the number of selectable $k$-card subsets is
\[
    R(x) = \prod_i \binom{c_i}{x_i}.
\]

\paragraph{Proof.}
For each mark $i$, we choose exactly $x_i$ of the $c_i$ cards having that mark. The choices are
independent across marks, so the counts multiply. \qed

\paragraph{Lemma 3.}
For a fixed mark pattern $x$, the number of possible table states is
\[
    L(x) = (k-1)! \sum_{h : x_h > 0}
    \binom{c_h}{x_h - 1} \prod_{i \ne h} \binom{c_i}{x_i}.
\]

\paragraph{Proof.}
Fix the hidden mark $h$. Then the visible cards must contain $x_h-1$ cards of mark $h$ and $x_i$ cards
of every other mark $i$. The number of ways to choose those visible cards is exactly
\[
    \binom{c_h}{x_h - 1} \prod_{i \ne h} \binom{c_i}{x_i}.
\]
After choosing them, the assistant may arrange them in any of $(k-1)!$ orders. Summing over every
possible hidden mark gives the formula. \qed

\paragraph{Lemma 4.}
For a fixed mark pattern $x$, the maximum number of selected sets that can be decoded correctly is
\[
    \min(L(x), R(x)).
\]

\paragraph{Proof.}
Model the situation as the bipartite graph described above: right vertices are selected $k$-sets of type
$x$, left vertices are table states of type $x$, and an edge means that the assistant can realize that
table state from that selected set.

Choosing a deterministic optimal strategy is exactly choosing distinct table states for as many selected
sets as possible, which is a matching. By the symmetry of this graph inside the fixed pattern class, Hall's
theorem guarantees that the smaller side can be saturated, so the maximum matching size is
\[
    \min(L(x), R(x)).
\]
\qed

\paragraph{Lemma 5.}
Conditioned on mark pattern $x$, the optimal success probability equals
\[
    \min\!\left(1,\,
    (k-1)! \sum_h \frac{x_h}{c_h - x_h + 1}\right).
\]

\paragraph{Proof.}
By Lemma 4 the conditional success probability is
\[
    \frac{\min(L(x),R(x))}{R(x)} = \min\!\left(1,\frac{L(x)}{R(x)}\right).
\]
Using Lemma 3 and canceling the common factor $R(x)$ yields
\[
    \frac{L(x)}{R(x)}
    = (k-1)! \sum_h
    \frac{\binom{c_h}{x_h - 1}}{\binom{c_h}{x_h}}
    = (k-1)! \sum_h \frac{x_h}{c_h - x_h + 1}.
    \qedhere
\]

\paragraph{Theorem.}
The algorithm outputs the highest possible probability of guessing the hidden card correctly.

\paragraph{Proof.}
By Lemma 1, patterns can be treated independently. For each pattern, Lemma 5 gives the optimal
conditional success probability, and Lemma 2 gives the probability that this pattern occurs. Summing these
weighted contributions over all feasible patterns yields exactly the optimal overall success probability.
\qed

\section*{Complexity Analysis}

The number of states $(x_1,\dots,x_m)$ is the number of weak compositions of $k$ into $m$ parts, which is
small because $k,m \le 10$. For each state we do only $O(mk)$ arithmetic, so the running time is tiny and
the memory usage is $O(m)$.

\section*{Implementation Notes}

\begin{itemize}[leftmargin=*]
    \item We evaluate $\binom{N}{r}$ directly with a short multiplicative loop. Since $r \le 10$, this is
    both fast and numerically stable enough with long doubles.
    \item The final formula never needs huge exact integers; we work only with probabilities and ratios.
\end{itemize}

\end{document}
