\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red!70!black},
  numbers=left,
  numberstyle=\tiny\color{gray},
  breaklines=true,
  frame=single,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 2010 -- Quality of Living}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Summary}
Given an $R \times C$ grid where each cell has a unique quality rating from $1$ to $RC$, find an $H \times W$ subgrid whose median is minimized. The median of $HW$ values is the $\lceil HW/2 \rceil$-th smallest value.

\section{Solution}

\subsection{Algorithm: Binary Search on the Median}
Binary search on the answer $m$. For a candidate median $m$, define an indicator matrix:
\[
B[i][j] =
\begin{cases}
1 & \text{if } \text{grid}[i][j] \le m, \\
0 & \text{otherwise.}
\end{cases}
\]
Let $k = \lceil HW / 2 \rceil$. There exists an $H \times W$ subgrid with median $\le m$ if and only if some subgrid contains at least $k$ cells with values $\le m$, i.e., $\sum B[i][j] \ge k$ over that subgrid. This is checked via 2D prefix sums.

\subsection{Correctness}

\begin{lemma}
The predicate ``there exists an $H \times W$ subgrid with median $\le m$'' is monotone in $m$.
\end{lemma}
\begin{proof}
If the predicate holds for $m$, it holds for all $m' \ge m$, since increasing $m$ can only increase the count of values $\le m'$ in any subgrid. Hence binary search applies.
\end{proof}

\begin{theorem}
The smallest $m$ for which the predicate holds is the minimum achievable median.
\end{theorem}
\begin{proof}
At the transition point $m^*$, some subgrid has at least $k$ values $\le m^*$, so its median is $\le m^*$. For $m^* - 1$, no subgrid has $k$ values $\le m^* - 1$, so every subgrid has median $\ge m^*$.
\end{proof}

\subsection{Complexity}
\begin{itemize}
    \item \textbf{Time:} $O(RC \log(RC))$ --- $O(\log(RC))$ binary search iterations, each taking $O(RC)$ to build prefix sums and check all subgrids.
    \item \textbf{Space:} $O(RC)$.
\end{itemize}

\section{Implementation}

\begin{lstlisting}
#include <bits/stdc++.h>
using namespace std;

int main() {
    ios::sync_with_stdio(false);
    cin.tie(nullptr);

    int R, C, H, W;
    cin >> R >> C >> H >> W;

    vector<vector<int>> grid(R, vector<int>(C));
    for (int i = 0; i < R; i++)
        for (int j = 0; j < C; j++)
            cin >> grid[i][j];

    int medianPos = (H * W + 1) / 2; // 1-indexed rank of the median

    auto check = [&](int m) -> bool {
        // Build prefix sum of indicator (grid[i][j] <= m)
        vector<vector<int>> ps(R + 1, vector<int>(C + 1, 0));
        for (int i = 0; i < R; i++)
            for (int j = 0; j < C; j++)
                ps[i+1][j+1] = (grid[i][j] <= m ? 1 : 0)
                    + ps[i][j+1] + ps[i+1][j] - ps[i][j];

        // Check all H x W subgrids
        for (int i = H; i <= R; i++)
            for (int j = W; j <= C; j++) {
                int cnt = ps[i][j] - ps[i-H][j] - ps[i][j-W] + ps[i-H][j-W];
                if (cnt >= medianPos) return true;
            }
        return false;
    };

    int lo = 1, hi = R * C, ans = hi;
    while (lo <= hi) {
        int mid = (lo + hi) / 2;
        if (check(mid)) {
            ans = mid;
            hi = mid - 1;
        } else {
            lo = mid + 1;
        }
    }

    cout << ans << "\n";
    return 0;
}
\end{lstlisting}

\end{document}
