\section*{Overview}

Minimum spanning tree is the weighted-graph tool I expect to need whenever the task says ``connect everything as
cheaply as possible'' without singling out one source or sink. In contests, Kruskal plus DSU is the version I reach for
first because it is short, stable, and easy to adapt.

\section*{When to Use It}

Use MST ideas when:

\begin{itemize}
  \item the graph is undirected and weighted,
  \item the goal is to connect all vertices with minimum total edge weight,
  \item the problem hides a connectivity threshold or ``bottleneck'' argument.
\end{itemize}

\section*{Core Idea}

A spanning tree uses exactly \(n - 1\) edges and connects every vertex. The minimum spanning tree is the spanning tree
with the smallest possible total weight.

Kruskal's algorithm sorts all edges by weight and scans from lightest to heaviest, adding an edge only when it connects
two different DSU components.

\section*{Key Insight}

The algorithm is safe because of the cut property: for any cut of the graph, the lightest edge crossing that cut can
belong to some MST. Kruskal keeps using that fact globally by always taking the cheapest edge that does not create a
cycle.

\section*{Operations / Main Technique}

\begin{itemize}
  \item sort edges by weight,
  \item maintain DSU components,
  \item add an edge iff its endpoints are currently in different components,
  \item stop after selecting \(n - 1\) edges.
\end{itemize}

\paragraph{Worked example.}
If the next lightest edge joins two vertices already connected through previously chosen edges, adding it would only
create a cycle and cannot improve the tree, so Kruskal skips it.

\section*{Correctness Intuition}

Whenever Kruskal picks the cheapest edge that joins two components, there exists some MST that includes it. Repeating
that exchange argument edge by edge shows that the whole constructed tree can be part of an MST, and once \(n - 1\)
edges are chosen, the tree is minimum.

\section*{Complexity Analysis}

\begin{itemize}
  \item sorting edges: \(O(m \log m)\),
  \item DSU operations: almost linear,
  \item overall: \(O(m \log m)\).
\end{itemize}

\section*{Implementation}

The reference code returns both:

\begin{itemize}
  \item the total weight,
  \item the list of chosen edges.
\end{itemize}

That is enough both for plain MST tasks and for follow-up tree processing.

\section*{Common Pitfalls}

\begin{itemize}
  \item Forgetting that MST is for undirected graphs.
  \item Assuming an MST exists even when the graph is disconnected.
  \item Using 32-bit sums when edge weights and \(n\) are large.
  \item Reaching for MST when the statement is actually about shortest paths from one source.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Prim's algorithm for dense graphs or adjacency-matrix settings.
  \item Maximum spanning tree by reversing the comparator.
  \item Second-best MST and replacement-edge arguments.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Standard minimum spanning tree construction.
  \item Minimum bottleneck connection problems.
  \item Build the MST first, then answer queries on the resulting tree.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/graph/mst_kruskal.html}{cp-algorithms: Minimum Spanning Tree - Kruskal}.
  \item \href{https://codeforces.com/catalog?locale=en}{Codeforces Catalog}.
  \item \href{https://usaco.guide/CPH.pdf}{Competitive Programmer's Handbook}.
\end{itemize}
