\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}

\title{ICPC World Finals 2023\\I. Waterworld}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Problem Summary}

The planet is partitioned into $n$ horizontal bands and $m$ equal rotation slices. For each band-slice
cell we are given the percentage covered by water, and we must compute the percentage of the whole
planet covered by water.

\section*{Key Observations}

\begin{itemize}[leftmargin=*]
    \item On a sphere of radius $R$, the surface area of a horizontal band of height $h$ is $2\pi Rh$.
    \item Therefore, in the cylindrical equal-area projection, equal vertical heights correspond to equal
    surface areas.
    \item Each input row describes one horizontal band of equal height, and each input column describes one
    rotation slice of equal angle. So every matrix entry corresponds to exactly the same surface area.
\end{itemize}

\section*{Algorithm}

\begin{enumerate}[leftmargin=*]
    \item Read all $nm$ percentages.
    \item Compute their arithmetic mean.
    \item Output that mean.
\end{enumerate}

\section*{Correctness Proof}

We prove that the algorithm returns the correct answer.

\paragraph{Lemma 1.}
All $nm$ cells represented by the input have equal surface area on the sphere.

\paragraph{Proof.}
Each row has the same height, so by the equal-area formula for spherical bands each row occupies the
same total surface area as every other row. Each column splits a row into equal angles, so it also splits
that row into equal areas. Hence every row-column cell has the same surface area. \qed

\paragraph{Theorem.}
The algorithm outputs the true percentage of the planet covered by water.

\paragraph{Proof.}
By Lemma 1, the whole planet is partitioned into $nm$ equal-area cells, and the input value of each cell is
its water percentage. The global water percentage is therefore the average of those $nm$ percentages,
which is exactly what the algorithm computes. \qed

\section*{Complexity Analysis}

The algorithm reads the matrix once, so the running time is $O(nm)$ and the memory usage is $O(1)$
besides the input values being streamed.

\section*{Implementation Notes}

\begin{itemize}[leftmargin=*]
    \item Using floating point only for the final average is enough; the input values themselves are integers.
\end{itemize}

\end{document}
