\section*{Overview}

Bridges and articulation points identify the fragile parts of an undirected graph. A bridge is an edge whose removal
disconnects the graph. An articulation point is a vertex whose removal does the same.

\section*{When to Use It}

Use these routines when:

\begin{itemize}
  \item the problem asks which roads, links, or nodes are critical,
  \item the graph is undirected and connectivity under deletions matters,
  \item biconnected structure is more relevant than shortest paths.
\end{itemize}

\section*{Core Idea}

Run DFS and compute:

\begin{itemize}
  \item \(\texttt{tin[v]}\): the DFS entry time,
  \item \(\texttt{low[v]}\): the smallest entry time reachable from \(v\)'s subtree using tree edges and at most one back edge.
\end{itemize}

These values tell whether a subtree can reconnect to an ancestor without using its parent edge.

\section*{Key Insight}

If a child subtree of \(v\) cannot reach an ancestor of \(v\), then the edge to that child is a bridge. The same
structural failure also tells when \(v\) separates that subtree from the rest of the graph.

\section*{Operations / Main Technique}

\begin{itemize}
  \item DFS with parent tracking,
  \item update \(\texttt{low}\) from tree children and back edges,
  \item test bridge condition \(\texttt{low[to] > tin[v]}\),
  \item test articulation condition with the root and non-root cases separated.
\end{itemize}

\paragraph{Worked example.}
If child \(\texttt{to}\) of \(v\) has \(\texttt{low[to] > tin[v]}\), then the subtree rooted at \(\texttt{to}\) has no
back edge reaching \(v\) or above. So removing edge \((v, to)\) disconnects that subtree.

\section*{Correctness Intuition}

The lowlink value summarizes the highest ancestor reachable from the subtree. If that reach does not climb above the
parent edge, then the parent edge is the only connection upward. That is exactly the bridge condition. The articulation
condition is the vertex version of the same idea.

\section*{Complexity Analysis}

With adjacency lists, the algorithm runs in \(O(n + m)\) time and \(O(n)\) extra memory.

\section*{Implementation}

The sample collects both:

\begin{itemize}
  \item all bridges,
  \item all articulation points.
\end{itemize}

It assumes a simple undirected graph. Multiedges need a little extra care in the parent-edge handling.

\section*{Common Pitfalls}

\begin{itemize}
  \item Forgetting the special root case for articulation points.
  \item Mishandling multiedges or self-loops.
  \item Updating \(\texttt{low}\) with the wrong endpoint on back edges.
  \item Treating the algorithm as if it worked unchanged on directed graphs.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Edge-biconnected and vertex-biconnected components.
  \item Bridge tree after contracting 2-edge-connected components.
  \item Offline connectivity reasoning with lowlink as a subroutine.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item List all bridges in a network.
  \item Count articulation points.
  \item Contract components after removing bridges and run tree DP on the result.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/graph/bridge-searching.html}{cp-algorithms: Finding Bridges in a Graph}.
  \item \href{https://cp-algorithms.com/graph/cutpoints.html}{cp-algorithms: Finding Articulation Points in a Graph}.
  \item \href{https://usaco.guide/adv/BCC-2CC}{USACO Guide: Biconnected Components}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
