\section*{Overview}

Digit DP is the technique for counting integers in a huge range when the property depends on the digits themselves.
Instead of iterating through every number up to \(10^{18}\), it builds the answer one digit at a time while
remembering only the information that still matters for the unfinished suffix.

It is one of those patterns that looks mysterious the first few times, but becomes routine once the three standard
state flags are internalized.

\section*{When to Use It}

Use it when:

\begin{itemize}
  \item the upper bound is enormous, so brute force is impossible,
  \item validity depends on decimal digits, digit sums, remainders, adjacency, masks, or automaton states,
  \item the natural range answer is \(\texttt{solve}(R) - \texttt{solve}(L - 1)\).
\end{itemize}

If the property has nothing to do with the decimal expansion, digit DP is usually forcing the wrong viewpoint.

\section*{Core Idea}

Write the upper bound \(N\) as a string and process digits from left to right.

The classic state usually contains:

\begin{itemize}
  \item \textbf{position}: which digit are we filling,
  \item \textbf{tight}: whether the prefix built so far is exactly equal to the prefix of \(N\),
  \item \textbf{started}: whether we have already placed a non-leading-zero digit,
  \item one problem-specific state such as digit sum, remainder, previous digit, or automaton node.
\end{itemize}

\section*{Key Insight}

The two flags that make digit DP feel natural are:

\begin{itemize}
  \item \textbf{tight}: while this is true, the current digit cannot exceed the corresponding digit of the limit,
  \item \textbf{started}: leading zeros should usually not affect states such as digit sum or ``previous digit''.
\end{itemize}

Once \texttt{tight} becomes false, the rest of the digits are unconstrained. That is why only non-tight states are
normally memoized.

\section*{Operations / Main Technique}

The standard recipe is:

\begin{itemize}
  \item define \(\texttt{solve}(N)\),
  \item run a DFS over digits with memoization,
  \item memoize only states with \(\texttt{tight = false}\),
  \item answer a range by subtraction.
\end{itemize}

\paragraph{Worked example.}
To count numbers in \([0, N]\) with digit sum \(= s\), the problem-specific state is just the sum accumulated so far.
Each transition chooses the next digit, updates the sum, updates \texttt{tight}, and decides whether the number has
really started.

\paragraph{Automaton-like states.}
In harder problems, the state can be:

\begin{itemize}
  \item remainder modulo \(m\),
  \item last digit,
  \item bitmask of used digits,
  \item position in a small string-matching automaton.
\end{itemize}

\section*{Correctness Intuition}

Every integer in \([0, N]\) corresponds to exactly one root-to-leaf path through the DP states: the path that writes
its decimal digits from left to right.

The \texttt{tight} flag prevents paths from exceeding \(N\). The problem-specific state remembers exactly the
information needed to decide validity once the last digit is placed. Two non-tight states with the same
\((position, started, problem\_state)\) have identical futures, so memoization is sound.

\section*{Complexity Analysis}

If the problem-specific state space has size \(S\), then the rough bound is:

\begin{itemize}
  \item time: \(O(\text{digits} \cdot S \cdot 10)\),
  \item memory: \(O(\text{digits} \cdot S)\).
\end{itemize}

For 64-bit bounds, the number of digits is tiny, so the real bottleneck is almost always the extra state dimension.

\section*{Implementation}

The reference code implements a clean teaching example:

\begin{itemize}
  \item count numbers whose digit sum is exactly a target,
  \item expose both \(\texttt{count\_exact\_sum(limit, sum)}\) and range subtraction,
  \item memoize only non-tight states.
\end{itemize}

That code is intentionally small, because once the \texttt{tight}/\texttt{started} logic is right, the rest of the
adaptation is usually in the problem-specific state.

\section*{Common Pitfalls}

\begin{itemize}
  \item Forgetting the \texttt{started} flag and accidentally counting leading zeros as real digits.
  \item Memoizing tight states even though they depend on the exact limit prefix.
  \item Returning the wrong base case for the number \(0\).
  \item Letting the custom state explode so the DP becomes too large.
  \item Writing \(\texttt{solve}(N)\) correctly but forgetting to subtract \(\texttt{solve}(L - 1)\).
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item digit DP with remainders modulo \(m\),
  \item no-adjacent-equal-digit constraints,
  \item bitmask digit DP,
  \item automaton digit DP for forbidden substrings,
  \item optimization where \texttt{started} is folded into the custom state when convenient.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Count numbers with a fixed digit sum.
  \item Count numbers with no adjacent equal digits.
  \item Range counting where validity depends on both digit remainder and local adjacency.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://usaco.guide/gold/digit-dp?lang=cpp}{USACO Guide: Digit DP}.
  \item \href{https://usaco.guide/PAPS.pdf}{Principles of Algorithmic Problem Solving}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
