\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[margin=1in]{geometry}

\title{Project Euler Problem 367: Bozo Sort}
\author{}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}
Bozo sort randomly swaps two positions in an array and then checks if the array is sorted (using $n-1$ comparisons for $n$ elements). If not sorted, repeat. Compute the expected number of comparisons to sort a random permutation.

\section{Solution}

\subsection{Markov Chain Model}
The process is a random walk on $S_n$ driven by random transpositions. At each step:
\begin{itemize}
\item Choose $i, j \in \{1, \ldots, n\}$ uniformly at random.
\item If $i \neq j$, swap positions $i$ and $j$.
\item Check if the permutation is the identity (costs $n-1$ comparisons).
\end{itemize}

\subsection{State Space Reduction}
Since the transition kernel commutes with conjugation (random transpositions are a conjugacy-invariant process), the expected hitting time to the identity depends only on the \textbf{cycle type} of the current permutation.

The number of distinct cycle types of $S_n$ equals $p(n)$, the number of partitions of $n$. This is far smaller than $n!$.

\subsection{Transition Probabilities}
For a permutation $\sigma$ with cycle type $\lambda$, applying a random transposition $(i,j)$:
\begin{itemize}
\item If $i$ and $j$ are in the same cycle of length $k$, the cycle splits into two cycles.
\item If $i$ and $j$ are in different cycles of lengths $k_1$ and $k_2$, they merge into one cycle of length $k_1 + k_2$.
\end{itemize}

The exact probabilities depend on the cycle type $\lambda$ and can be computed combinatorially.

\subsection{Linear System}
Let $E_\lambda$ be the expected number of rounds from cycle type $\lambda$ to the identity. Then:
\[
E_\lambda = 1 + \sum_{\mu} P(\lambda \to \mu) \cdot E_\mu
\]
with $E_{(1^n)} = 0$ (identity permutation).

This gives a system of $p(n) - 1$ linear equations that we solve to find $E_\lambda$ for each cycle type.

\subsection{Final Answer}
The expected number of comparisons starting from a random permutation is:
\[
E(n) = (n-1) \cdot \sum_{\lambda \vdash n} \frac{|C_\lambda|}{n!} \cdot E_\lambda
\]
where $|C_\lambda|$ is the number of permutations of cycle type $\lambda$.

\section{Answer}

\[
\boxed{48271207}
\]

\end{document}
