\section*{Overview}

The sieve is the standard way to turn repeated prime questions into preprocessing. Instead of testing each number from
scratch, I mark composite numbers once and reuse that work everywhere else.

\section*{When to Use It}

Use a sieve when:

\begin{itemize}
  \item many primality or factorization queries are asked up to a common bound,
  \item you need all primes up to \(n\),
  \item multiplicative preprocessing starts from prime information.
\end{itemize}

\section*{Core Idea}

Initialize every number as potentially prime, then iterate through the range. When a prime \(p\) is found, mark its
multiples as composite. Each composite is crossed out because it has a smaller prime divisor.

\section*{Key Insight}

The sieve wins because composite numbers are processed by structure rather than individually. After that, primality and
smallest-prime-factor queries become array lookups.

\section*{Operations / Main Technique}

\begin{itemize}
  \item build a boolean primality table,
  \item optionally record the smallest prime factor of each number,
  \item factor numbers quickly by repeatedly dividing by the stored smallest prime factor.
\end{itemize}

\paragraph{Worked example.}
Once \(2\) is known prime, mark \(4, 6, 8, \ldots\). Once \(3\) is known prime, mark \(9, 12, 15, \ldots\). Numbers
that survive all smaller prime marks are prime.

\section*{Correctness Intuition}

Every composite number has a smallest prime divisor. When that prime is processed, the composite is marked. A prime is
never marked by a smaller factor because none exists.

\section*{Complexity Analysis}

\begin{itemize}
  \item classic sieve: \(O(n \log \log n)\),
  \item linear smallest-prime-factor sieve: \(O(n)\),
  \item one factorization using SPF: \(O(\log n)\) in practice.
\end{itemize}

\section*{Implementation}

The reference code uses the linear sieve pattern because it naturally produces:

\begin{itemize}
  \item a list of primes,
  \item a smallest-prime-factor array,
  \item fast factorization support afterward.
\end{itemize}

\section*{Common Pitfalls}

\begin{itemize}
  \item Choosing a bound that is too small for later queries.
  \item Forgetting that the sieve is preprocessing, so it only helps if many queries share one limit.
  \item Using too much memory when the bound is very large.
  \item Confusing primality with smallest prime factor \(= 0\) or \(= i\) depending on the implementation.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Segment sieve for a large interval \([L, R]\).
  \item Euler phi and Mobius preprocessing.
  \item Prime-counting prefix arrays on top of the primality table.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Answer many primality queries.
  \item Factor all array values up to a fixed limit.
  \item Precompute multiplicative functions for many numbers.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/algebra/sieve-of-eratosthenes.html}{cp-algorithms: Sieve of Eratosthenes}.
  \item \href{https://usaco.guide/gold/prime-factorization}{USACO Guide: Prime Factorization and Sieve}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
