\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
\title{Problem 473: Phigital Number Base}
\author{Project Euler Solutions}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}
Let $\varphi = \frac{1+\sqrt{5}}{2}$ be the golden ratio. Every positive integer has a unique ``phigital'' representation as a sum of non-consecutive powers of $\varphi$ (analogous to the Zeckendorf representation). A phigital representation is \emph{palindromic} if the digit string reads the same forwards and backwards, ignoring the radix point.

Find the sum of all positive integers not exceeding $10^{10}$ whose phigital representation is palindromic.

\section{Mathematical Analysis}
\subsection{Base-$\varphi$ Representation}
The golden ratio satisfies $\varphi^2 = \varphi + 1$. Every positive integer $n$ can be written uniquely as:
\[
n = \sum_{i} d_i \varphi^i, \quad d_i \in \{0,1\}, \quad d_i d_{i+1} = 0 \;\forall\, i
\]
with finitely many nonzero digits (extending to negative exponents).

\subsection{Palindromic Structure}
For a representation with digits $d_k d_{k-1} \cdots d_1 d_0 \,.\, d_0' d_1' \cdots d_k'$, palindromicity means $d_i = d_i'$ for all $i$. Thus a palindromic number has the form:
\[
n = \sum_{i \in S} \left(\varphi^i + \varphi^{-(i+1)}\right)
\]
where $S$ is a set of non-negative integers with no two consecutive elements.

\subsection{Integrality}
Since $\varphi^n + \psi^n = L_n$ (the $n$-th Lucas number) and $|\psi| < 1$, the palindromic sums naturally produce values close to integers. The non-consecutive constraint ensures uniqueness.

\section{Algorithm}
\begin{enumerate}
    \item Precompute $\varphi^i$ for $i = 0, \ldots, 48$ using extended-precision arithmetic.
    \item Use depth-first search over bit positions $k, k-1, \ldots, 0$, deciding whether to activate each palindromic pair.
    \item Enforce the non-consecutive constraint: if position $i$ is active, skip position $i-1$.
    \item Prune branches where the accumulated value exceeds $10^{10}$.
    \item Collect all integer-valued results and sum them.
\end{enumerate}

\section{Complexity Analysis}
\begin{itemize}
    \item \textbf{Time:} $O(\varphi^{k/2})$ where $k \approx 47$, yielding roughly $10^5$ candidates due to the Fibonacci-like branching of the non-consecutive constraint.
    \item \textbf{Space:} $O(k)$ for the recursion stack and precomputed powers.
\end{itemize}

\section{Answer}

\[
\boxed{35856681704365}
\]

\end{document}
