\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
\title{Problem 913: Balanced Ternary Representation}
\author{Project Euler Solutions}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}
In balanced ternary, digits are $\{-1, 0, 1\}$ (written as $T, 0, 1$). Every integer has a unique balanced ternary representation. Find the sum of all positive integers $n \leq 10^6$ whose balanced ternary representation contains no zero digit.

\section{Mathematical Analysis}
A number in balanced ternary uses digits $d_i \in \{-1, 0, 1\}$ so that $n = \sum_i d_i \cdot 3^i$. We need representations with only digits $\pm 1$.

Numbers expressible as $\sum_{i=0}^{k} \epsilon_i 3^i$ where $\epsilon_i \in \{-1, 1\}$ form a specific subset. For $k$ digits, there are $2^{k+1}$ such numbers (half positive, half negative, by symmetry).

\section{Derivation}
For a $k$-digit balanced ternary number (digits $d_0, \ldots, d_k$ each $\pm 1$), the value is $\sum_{i=0}^{k} d_i \cdot 3^i$.

By symmetry, for each positive such number $n$, $-n$ is also representable. The positive numbers have $d_k = 1$.

We enumerate all $2^k$ choices for $(d_0,\ldots,d_{k-1})$ with $d_k=1$ and collect those with value in $[1, 10^6]$.

Max $k$: $3^k \leq 10^6 + (3^k - 1)/2$ gives $k \leq 12$ (since $3^{12} = 531441$, and $3^{13} = 1594323 > 10^6$).

\section{Proof of Correctness}
Each balanced ternary representation is unique. We generate all values with no-zero digits and filter for $[1, 10^6]$.

\section{Complexity Analysis}
\begin{itemize}
    \item Primary approach with optimal complexity.
    \item Brute-force verification for small inputs.
\end{itemize}

\section{Answer}

\[
\boxed{2101925115560555020}
\]

\end{document}
