\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[margin=1in]{geometry}

\title{Problem 189: Tri-colouring a Triangular Grid}
\author{Project Euler Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

A triangular grid of side 8 is divided into $8^2 = 64$ small triangles. Each triangle must be coloured with one of 3 colours such that no two triangles sharing an edge have the same colour. How many valid colourings exist?

\section{Mathematical Analysis}

\subsection{Grid Structure}

Row $k$ ($1 \le k \le 8$) contains $2k-1$ small triangles: $k$ upward-pointing and $k-1$ downward-pointing. Triangles within a row alternate between upward and downward.

\subsection{Adjacency}

Two triangles are adjacent if they share an edge. The adjacencies are:
\begin{itemize}
    \item Within a row: consecutive triangles share a vertical edge.
    \item Between rows: each downward triangle in row $k+1$ shares its top edge with the upward triangle at the corresponding position in row $k$.
\end{itemize}

\subsection{Dynamic Programming}

We process rows top to bottom. The \textbf{state} for row $k$ is the tuple $(c_1, c_2, \ldots, c_k)$ of colours assigned to the $k$ upward-pointing triangles.

\textbf{Transition from row $k$ to row $k+1$:}

Given the colours of upward triangles in row $k$, we must:
\begin{enumerate}
    \item Choose colours for the $k+1$ upward triangles of row $k+1$.
    \item Choose colours for the $k$ downward triangles of row $k+1$.
    \item Ensure all adjacency constraints are satisfied.
\end{enumerate}

Each downward triangle $j$ in row $k+1$ must differ from:
\begin{itemize}
    \item Upward triangle $j$ in row $k$ (top neighbor),
    \item Upward triangle $j$ in row $k+1$ (left neighbor),
    \item Upward triangle $j+1$ in row $k+1$ (right neighbor).
\end{itemize}

Additionally, consecutive upward triangles in row $k+1$ must not share the same colour as any intervening downward triangle (but these constraints are already captured by the downward triangle constraints combined with the within-row adjacency).

\subsection{State Space}

The number of states per row is at most $3^k$ for row $k$. Since $k \le 8$, this is at most $3^8 = 6561$, making the DP entirely feasible.

\section{Complexity}

\begin{itemize}
    \item \textbf{Time}: $O(n \cdot 3^{2n})$ for the transition computation, where $n = 8$.
    \item \textbf{Space}: $O(3^n)$ for DP state storage.
\end{itemize}

\section{Answer}

\[
\boxed{10834893628237824}
\]

\end{document}
