\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,booktabs,algorithm2e}
\usepackage[margin=1in]{geometry}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\title{Problem 871: Drifting Digits}
\author{Project Euler Solutions}
\date{}
\begin{document}
\maketitle

\section{Problem Statement}

Consider the \emph{reverse-and-add} process: given a positive integer $n$, compute $n + \operatorname{rev}(n)$ where $\operatorname{rev}(n)$ reverses the base-10 digits of $n$. Iterate until a palindrome is reached. The problem asks for a specific aggregate derived from this iteration applied across a parametrized family of starting values.

\section{Mathematical Foundation}

\begin{definition}[Reverse-and-Add]
Let $n = \sum_{i=0}^{d-1} a_i \cdot 10^i$ with $a_{d-1} \neq 0$. Define $\operatorname{rev}(n) = \sum_{i=0}^{d-1} a_{d-1-i} \cdot 10^i$ and the map $T(n) = n + \operatorname{rev}(n)$.
\end{definition}

\begin{definition}[Lychrel Number]
A natural number $n$ is a \emph{Lychrel number} if the sequence $(T^{(k)}(n))_{k \ge 0}$ never produces a palindrome. The existence of such numbers in base~10 remains open; 196 is the smallest candidate.
\end{definition}

\begin{theorem}[Carry Propagation]
Let $n$ have digits $a_0, a_1, \ldots, a_{d-1}$ (least significant first). In $n + \operatorname{rev}(n)$, position~$i$ has digit
\[
s_i \equiv a_i + a_{d-1-i} + c_i \pmod{10},
\]
where $c_0 = 0$ and $c_{i+1} = \lfloor (a_i + a_{d-1-i} + c_i)/10 \rfloor$. The result is a palindrome iff $s_i \bmod 10 = s_{d'-1-i} \bmod 10$ for all valid~$i$, where $d'$ is the digit count of $T(n)$.
\end{theorem}

\begin{proof}
Standard base-10 addition: position $i$ of $n$ contributes $a_i$; position $i$ of $\operatorname{rev}(n)$ contributes $a_{d-1-i}$. The carry recurrence is the usual one. The palindrome condition is that the resulting digit sequence reads identically forwards and backwards.
\end{proof}

\begin{lemma}[Carry-Free Case]
If $a_i + a_{d-1-i} \le 9$ for all $i$, then $T(n)$ is a palindrome.
\end{lemma}

\begin{proof}
Without carries, $s_i = a_i + a_{d-1-i} = s_{d-1-i}$, so the digit sequence is symmetric.
\end{proof}

\begin{theorem}[Digit Growth Bound]
The digit count of $T^{(k)}(n)$ is at most $d + k$, where $d$ is the initial digit count.
\end{theorem}

\begin{proof}
Each application of $T$ satisfies $T(n) \le 2 \cdot 10^{d'} - 2$ where $d'$ is the digit count of $n$, so the digit count increases by at most~1 per step.
\end{proof}

\section{Algorithm}

\begin{algorithm}[H]
\SetAlgoLined
\KwIn{Problem parameters}
\KwOut{Answer}
\For{each starting value $n$}{
    \For{$\text{step} = 1$ \KwTo $\text{MAX\_ITER}$}{
        $r \leftarrow \operatorname{rev}(n)$\;
        $n \leftarrow n + r$\;
        \If{$n$ is a palindrome}{
            record $(n, \text{step})$; \textbf{break}\;
        }
    }
}
Aggregate results and \Return answer\;
\end{algorithm}

\section{Complexity Analysis}

\begin{itemize}
\item \textbf{Time:} $O(S \cdot d)$ per starting value where $S$ = steps and $d$ = max digit count. Since $d \le d_0 + S$, the per-value cost is $O(S^2)$ with big-integer arithmetic.
\item \textbf{Space:} $O(d)$ for the current number.
\end{itemize}

\section{Answer}

\[
\boxed{2848790}
\]

\end{document}
