\section*{Overview}

Euler tour technique is the standard way to turn tree structure into array intervals. Once each subtree becomes one
contiguous segment, the rest of the problem often becomes a Fenwick-tree or segment-tree task instead of a tree task.

\section*{When to Use It}

Use Euler tour technique when:

\begin{itemize}
  \item queries are about subtrees of a rooted tree,
  \item you want to map tree nodes to an array,
  \item another array-based structure already solves the interval version.
\end{itemize}

\section*{Core Idea}

Run DFS and record the entry time \(\texttt{tin[v]}\) of each node. If nodes are appended to an order array at entry
time, then the subtree of \(v\) occupies one contiguous interval:
\[
[\texttt{tin[v]}, \texttt{tout[v]}].
\]

\section*{Key Insight}

DFS enters a subtree, finishes all of it, and only then leaves. That nesting property is exactly why every rooted
subtree becomes a single interval in entry order.

\section*{Operations / Main Technique}

\begin{itemize}
  \item root the tree,
  \item DFS once to compute \(\texttt{tin}\), \(\texttt{tout}\), and the flattened order,
  \item convert a subtree query into one array interval,
  \item answer with Fenwick or segment tree if updates are also present.
\end{itemize}

\paragraph{Worked example.}
If node \(v\) gets entry time \(7\) and its subtree ends at time \(12\), then ``sum over subtree \(v\)'' becomes ``sum
over interval \([7,12]\)'' on the flattened array.

\section*{Correctness Intuition}

During DFS, every node in the subtree of \(v\) is visited after entering \(v\) and before leaving it. No outside node
can appear inside that interval, so the entry-order slice exactly matches the subtree.

\section*{Complexity Analysis}

The flattening DFS is \(O(n)\). After that, the complexity is whatever interval structure is used on top of the order
array.

\section*{Implementation}

The code provides a reusable flattener with:

\begin{itemize}
  \item \texttt{tin},
  \item \texttt{tout},
  \item \texttt{order},
  \item subtree interval accessors.
\end{itemize}

\section*{Common Pitfalls}

\begin{itemize}
  \item Forgetting to root the tree consistently.
  \item Confusing node times with edge times in edge-weighted tasks.
  \item Assuming the same flattening also makes arbitrary paths contiguous. It usually does not.
  \item Mixing inclusive and half-open interval conventions.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Full Euler tour with entry and exit duplicates for LCA/RMQ style reductions.
  \item Subtree add and subtree sum with a Fenwick or segment tree.
  \item Mo on trees by combining an Euler tour with offline query ordering.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Subtree sum with point updates.
  \item Count colors or active markers inside subtrees.
  \item Convert a tree DP precomputation into subtree interval updates.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://codeforces.com/catalog?locale=en}{Codeforces Catalog}.
  \item \href{https://usaco.guide/plat/sqrt}{USACO Guide modules using tree flattening and Euler order}.
  \item \href{https://github.com/kth-competitive-programming/kactl}{KACTL}.
\end{itemize}
