\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}

\title{ICPC World Finals 2023\\A. Riddle of the Sphinx}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Problem Summary}

There are three unknown nonnegative integers: the numbers of legs of an axex, a basilisk, and a
centaur. We may ask exactly five linear questions of the form
\[
ax + by + cz = r,
\]
but one of the five answers may be a lie. We must still determine the unique triple $(x,y,z)$.

\section*{Key Observations}

\begin{itemize}[leftmargin=*]
    \item Each question gives one linear equation in the three unknowns.
    \item If we choose five query vectors such that \emph{any} three of them are linearly independent, then
    after discarding the one lie the remaining truthful answers determine a unique solution.
    \item A convenient fixed choice is
    \[
    (1,0,0),\ (0,1,0),\ (0,0,1),\ (1,1,1),\ (1,2,3).
    \]
    Any three of these rows have nonzero determinant.
    \item Therefore, for each candidate lie position we can solve a $3 \times 3$ system and then verify the
    resulting triple against the four answers assumed to be truthful.
\end{itemize}

\section*{Algorithm}

\begin{enumerate}[leftmargin=*]
    \item Ask the five fixed questions above and read the five answers.
    \item For each index $i$, pretend the $i$th answer is the lie.
    \item Solve the linear system from any three of the other four equations with Cramer's rule.
    \item Reject the candidate if the system is singular, if any coordinate is negative, or if any of the four
    supposedly truthful equations is violated.
    \item Output the unique candidate that survives.
\end{enumerate}

\section*{Correctness Proof}

We prove that the algorithm returns the correct answer.

\paragraph{Lemma 1.}
For the chosen five query vectors, any three are linearly independent.

\paragraph{Proof.}
This is a direct determinant check. The first three rows are the identity matrix. Any triple containing at
least two of them is obviously independent, and the remaining mixed triples with $(1,1,1)$ and
$(1,2,3)$ also have nonzero determinant. Hence every triple is independent. \qed

\paragraph{Lemma 2.}
If we discard the actual lie, the algorithm reconstructs the true values $(x,y,z)$.

\paragraph{Proof.}
After removing the lie, the remaining four equations are all truthful. By Lemma 1, any three of their row
vectors are independent, so the corresponding $3 \times 3$ system has a unique solution. The true triple
satisfies those equations, so uniqueness implies that the solved triple is exactly the true one. It also
passes the verification against all four truthful equations. \qed

\paragraph{Theorem.}
The algorithm outputs the correct answer for every valid input.

\paragraph{Proof.}
By Lemma 2, when the algorithm guesses the actual lie position, it keeps the true triple. Now consider any
wrong lie position. Then at least one lie remains among the four equations checked during verification, so
the candidate cannot satisfy all four unless that ``lie'' happens to match the truthful value, which would
mean it was not a lie at all. Therefore every wrong guess is rejected, and the only surviving output is the
true triple. \qed

\section*{Complexity Analysis}

The number of questions and candidate lie positions is constant. The running time is $O(1)$ and the
memory usage is $O(1)$.

\section*{Implementation Notes}

\begin{itemize}[leftmargin=*]
    \item Because this is interactive, each question must be flushed immediately after printing. The
    implementation uses \texttt{endl}.
    \item All arithmetic easily fits in 64-bit signed integers.
\end{itemize}

\end{document}
