\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red!70!black},
  numbers=left,
  numberstyle=\tiny\color{gray},
  breaklines=true,
  frame=single,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 2013 -- Art Class}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Summary}
Given an image (2D grid of RGB pixels), classify it into one of four categories:
\begin{enumerate}
    \item \textbf{Class 1:} Modern/abstract art (few colors, large uniform regions)
    \item \textbf{Class 2:} Paintings with distinct objects (moderate variation)
    \item \textbf{Class 3:} Impressionist/textured paintings (high local variation)
    \item \textbf{Class 4:} Landscape/nature photographs (moderate to high detail)
\end{enumerate}

This is a heuristic/machine learning problem. No exact algorithm is specified; you must achieve a reasonable classification accuracy.

\section{Solution Approach}
Use simple image features:

\begin{enumerate}
    \item \textbf{Color variance:} Compute the variance of R, G, B channels globally. Low variance suggests Class 1 (uniform colors).
    \item \textbf{Edge density:} Compute the average absolute difference between adjacent pixels. Low edge density suggests large uniform regions (Class 1), very high suggests Class 3.
    \item \textbf{Green ratio:} High green content suggests landscapes (Class 4).
    \item \textbf{Saturation:} High saturation with high variance suggests impressionist (Class 3).
\end{enumerate}

\textbf{Decision rules:}
\begin{itemize}
    \item If edge density is very low and color variance is low: Class 1.
    \item If green ratio is high and edge density is moderate: Class 4.
    \item If edge density is very high: Class 3.
    \item Otherwise: Class 2.
\end{itemize}

\section{Complexity}
\begin{itemize}
    \item \textbf{Time:} $O(H \times W)$ (single pass over all pixels).
    \item \textbf{Space:} $O(H \times W)$
\end{itemize}

\section{C++ Solution}
\begin{lstlisting}
#include <bits/stdc++.h>
using namespace std;

int style(int H, int W, int R[500][500], int G[500][500], int B[500][500]){
    // Compute average color
    double avgR = 0, avgG = 0, avgB = 0;
    for(int i = 0; i < H; i++){
        for(int j = 0; j < W; j++){
            avgR += R[i][j];
            avgG += G[i][j];
            avgB += B[i][j];
        }
    }
    int total = H * W;
    avgR /= total; avgG /= total; avgB /= total;

    // Compute color variance
    double varR = 0, varG = 0, varB = 0;
    for(int i = 0; i < H; i++){
        for(int j = 0; j < W; j++){
            varR += (R[i][j] - avgR) * (R[i][j] - avgR);
            varG += (G[i][j] - avgG) * (G[i][j] - avgG);
            varB += (B[i][j] - avgB) * (B[i][j] - avgB);
        }
    }
    varR /= total; varG /= total; varB /= total;
    double totalVar = varR + varG + varB;

    // Compute edge density (average absolute difference with neighbors)
    double edgeSum = 0;
    int edgeCount = 0;
    for(int i = 0; i < H; i++){
        for(int j = 0; j < W - 1; j++){
            edgeSum += abs(R[i][j] - R[i][j+1]) + abs(G[i][j] - G[i][j+1])
                     + abs(B[i][j] - B[i][j+1]);
            edgeCount++;
        }
    }
    for(int i = 0; i < H - 1; i++){
        for(int j = 0; j < W; j++){
            edgeSum += abs(R[i][j] - R[i+1][j]) + abs(G[i][j] - G[i+1][j])
                     + abs(B[i][j] - B[i+1][j]);
            edgeCount++;
        }
    }
    double edgeDensity = edgeSum / edgeCount;

    // Green ratio
    double greenRatio = avgG / (avgR + avgG + avgB + 1);

    // Classification heuristic
    if(totalVar < 1000 && edgeDensity < 15){
        return 1; // Modern art: few colors, uniform regions
    }
    if(edgeDensity > 60){
        return 3; // Impressionist: very textured
    }
    if(greenRatio > 0.38 && totalVar > 2000){
        return 4; // Landscape: lots of green, varied
    }
    if(edgeDensity < 25 && totalVar < 3000){
        return 1;
    }
    if(edgeDensity > 40){
        return 3;
    }
    if(greenRatio > 0.35){
        return 4;
    }
    return 2; // Default: standard painting
}

int main(){
    int H, W;
    cin >> H >> W;
    int R[500][500], G[500][500], B[500][500];
    for(int i = 0; i < H; i++)
        for(int j = 0; j < W; j++)
            cin >> R[i][j] >> G[i][j] >> B[i][j];
    cout << style(H, W, R, G, B) << "\n";
    return 0;
}
\end{lstlisting}

\textbf{Note:} This is a heuristic solution. The thresholds were chosen empirically. More sophisticated approaches would use trained classifiers, but this demonstrates the feature-extraction approach suitable for a competition setting.

\end{document}
