\section*{Overview}

Meet-in-the-middle is what I use when \(2^n\) is too large but \(2^{n/2}\) is still comfortable. The trick is to split
the objects into two halves, enumerate each half independently, and then combine the partial results more intelligently
than a full brute force would.

\section*{When to Use It}

Use it when:

\begin{itemize}
  \item the direct search is exponential in \(n\),
  \item \(n\) is around \(35\) to \(45\), so \(2^{n/2}\) is still manageable,
  \item the objective can be expressed by combining one partial choice from the left half and one from the right half.
\end{itemize}

Classic patterns are subset sum, balanced partition, small-set knapsack, and ``find two partial states that fit
together'' problems.

\section*{Core Idea}

Split the items into two halves:
\[
L = \{a_0, \dots, a_{m-1}\},
\qquad
R = \{a_m, \dots, a_{n-1}\}.
\]
Enumerate all subset summaries of each half. Then combine the two lists by sorting, binary search, two pointers, or a
hash map depending on the objective.

\section*{Key Insight}

The exponential part is still there, but the exponent is cut in half. That is a huge change:
\[
2^{40} \text{ is impossible in contest time, but } 2^{20} + 2^{20} \text{ is completely normal.}
\]

\section*{Operations / Main Technique}

\begin{itemize}
  \item enumerate all subset sums or other half summaries,
  \item sort one side,
  \item for each value on the other side, find the best compatible partner.
\end{itemize}

\paragraph{Worked example.}
To find the largest subset sum at most \(S\), enumerate all sums on the left and right. Sort the right sums. For each
left sum \(x\), binary search the largest right sum \(y\) with \(x + y \le S\).

\section*{Correctness Intuition}

Every full subset is uniquely determined by its choice on the left half and its choice on the right half. So once both
half enumerations are complete, combining them explores the full search space without ever enumerating all \(2^n\)
states directly.

\section*{Complexity Analysis}

For \(n\) items, the standard subset-sum version runs in:
\begin{itemize}
  \item \(O(2^{n/2})\) to enumerate each half,
  \item \(O(2^{n/2} \log 2^{n/2})\) to sort and search,
  \item memory \(O(2^{n/2})\).
\end{itemize}

\section*{Implementation}

The reference code solves the ``maximum subset sum at most \(S\)'' template. That is the most reusable starting point,
and many other meet-in-the-middle problems are just different summaries or a different merge step.

\section*{Common Pitfalls}

\begin{itemize}
  \item Using meet-in-the-middle when \(n\) is already small enough for a simpler DP or brute force.
  \item Forgetting the memory cost of storing all half summaries.
  \item Generating all half states correctly but combining them with the wrong inequality or boundary handling.
  \item Missing dominance pruning when one half can be compressed further.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Split knapsack by weight or value summaries.
  \item Enumerate masks on one side and hashed states on the other.
  \item Use Pareto pruning when the summary has two dimensions.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Maximum subset sum not exceeding a limit.
  \item Closest subset sum to a target.
  \item Count subset pairs satisfying a constraint after sorting one half.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://codeforces.com/catalog?locale=en}{Codeforces Catalog}.
  \item \href{https://usaco.guide/CPH.pdf}{Competitive Programmer's Handbook}.
  \item \href{https://oi-wiki.org/search/mitm/}{OI Wiki: Meet in the Middle}.
\end{itemize}
