¥documentclass[12pt]{article}
¥usepackage{amsmath,amssymb,amsthm}
¥usepackage[margin=1in]{geometry}

¥newtheorem{theorem}{Theorem}
¥newtheorem{lemma}[theorem]{Lemma}

¥title{Project Euler Problem 501: Eight Divisors}
¥author{}
¥date{}

¥begin{document}
¥maketitle

¥section{Problem Statement}
Let $f(n)$ be the count of numbers not exceeding $n$ with exactly eight divisors.
Given: $f(100) = 10$, $f(1000) = 180$, $f(10^6) = 224427$. Find $f(10^{12})$.

¥section{Mathematical Foundation}

¥begin{theorem}[Divisor Count Formula]
Let $n = p_1^{a_1} p_2^{a_2} ¥cdots p_k^{a_k}$. Then $¥tau(n) = ¥prod_{i=1}^{k}(a_i + 1)$.
¥end{theorem}

¥begin{proof}
Each divisor of $n$ has the form $¥prod p_i^{b_i}$ with $0 ¥le b_i ¥le a_i$. There are $a_i + 1$ independent choices for each $b_i$, so by the multiplication principle, $¥tau(n) = ¥prod(a_i + 1)$.
¥end{proof}

¥begin{lemma}[Classification of $¥tau(n) = 8$]
A positive integer $n$ satisfies $¥tau(n) = 8$ if and only if $n$ takes one of the forms:
¥begin{enumerate}
    ¥item $n = p^7$ for some prime‾$p$,
    ¥item $n = p^3 q$ for distinct primes $p, q$,
    ¥item $n = pqr$ for distinct primes $p < q < r$.
¥end{enumerate}
¥end{lemma}

¥begin{proof}
We need $¥prod(a_i + 1) = 8 = 2^3$. The multiplicative partitions of‾8 into factors $¥ge 2$ are $¥{8¥}$, $¥{4, 2¥}$, and $¥{2, 2, 2¥}$, yielding the three forms above. These are clearly mutually exclusive and exhaust all possibilities.
¥end{proof}

¥begin{theorem}[Counting Formula]
Let $N = 10^{12}$ and $¥pi(x)$ denote the prime counting function. Then
¥[
f(N) = C_1 + C_2 + C_3
¥]
where:
¥begin{align}
C_1 &= ¥pi(N^{1/7}), ¥¥
C_2 &= ¥sum_{¥substack{p ¥text{ prime} ¥¥ p^3 < N}} ¥left[¥pi¥!¥left(¥frac{N}{p^3}¥right) - ¥mathbf{1}[p^4 ¥le N]¥right], ¥¥
C_3 &= ¥sum_{p ¥text{ prime}} ¥sum_{¥substack{q ¥text{ prime} ¥¥ p < q,¥; pq^2 ¥le N}} ¥left[¥pi¥!¥left(¥frac{N}{pq}¥right) - ¥pi(q)¥right].
¥end{align}
¥end{theorem}

¥begin{proof}
¥textbf{Case 1:} $n = p^7 ¥le N$ iff $p ¥le N^{1/7}$, giving count $¥pi(N^{1/7})$.

¥textbf{Case 2:} $n = p^3 q ¥le N$ with $p ¥ne q$. For each prime‾$p$ with $p^3 < N$, the number of valid primes $q ¥le N/p^3$ is $¥pi(N/p^3)$, minus‾1 when $q = p$ is itself valid (i.e., when $p^4 ¥le N$).

¥textbf{Case 3:} $n = pqr$ with $p < q < r$. For each pair $(p,q)$ with $p < q$, valid primes $r$ satisfy $q < r ¥le N/(pq)$, giving $¥pi(N/(pq)) - ¥pi(q)$ choices. We need $pq^2 < N$ for at least one valid‾$r$.

By Lemma‾1 the three cases are disjoint and exhaustive.
¥end{proof}

¥section{Algorithm}
¥begin{verbatim}
1. Sieve primes up to sqrt(N) = 10^6.
2. Implement pi(x) via Lucy Hedgehog / Meissel-Lehmer.
3. Case 1: C1 = pi(floor(N^(1/7))).
4. Case 2: For each prime p with p^3 < N:
     C2 += pi(N/p^3) - [p^4 <= N].
5. Case 3: For each prime p, for each prime q > p
     with p*q^2 <= N:
     C3 += pi(N/(p*q)) - pi(q).
6. Return C1 + C2 + C3.
¥end{verbatim}

¥section{Complexity Analysis}
¥begin{itemize}
    ¥item ¥textbf{Time:} $O(N^{2/3})$, dominated by the prime counting function.
    ¥item ¥textbf{Space:} $O(N^{1/2})$ for the sieve and $¥pi(x)$ tables.
¥end{itemize}

¥section{Answer}

¥[
¥boxed{197912312715}
¥]

¥end{document}
