\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red!70!black},
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=8pt,
  frame=single,
  breaklines=true,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 1995 -- Wires}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

Given $n$ points (terminals) in the plane, connect them all with wires of minimum total
length. Each wire connects exactly two terminals, and all terminals must be connected
(directly or indirectly).

This is the \textbf{Minimum Spanning Tree (MST)} problem on a complete graph where
edge weights are Euclidean distances.

\medskip
\noindent\textbf{Constraints:} $n \le 50$.

\section{Solution Approach}

\subsection{Minimum Spanning Tree}

Since we need to connect all points with minimum total wire length and any two points
can be directly connected, this is a standard MST problem.

\begin{enumerate}
  \item Compute all $\binom{n}{2}$ pairwise Euclidean distances.
  \item Apply Kruskal's algorithm: sort edges by weight and greedily add the shortest
        edge that does not create a cycle, using a Union-Find data structure.
  \item The MST gives the minimum total wiring length.
\end{enumerate}

\subsection{Why MST Is Optimal}

Among all spanning trees of the complete graph, the MST minimizes the total edge weight.
Since every spanning tree uses exactly $n-1$ edges and connects all nodes, the MST
achieves the minimum total wire length.

\textbf{Note:} A Steiner tree (which may introduce additional intermediate points) could
yield shorter total length, but the problem requires direct terminal-to-terminal connections.

\section{C++ Solution}

\begin{lstlisting}
#include <cstdio>
#include <cmath>
#include <algorithm>
#include <vector>
using namespace std;

struct Edge {
    double w;
    int u, v;
    bool operator<(const Edge& o) const { return w < o.w; }
};

int par[55], rnk[55];

int find(int x) {
    while (par[x] != x) x = par[x] = par[par[x]];
    return x;
}

bool unite(int a, int b) {
    a = find(a); b = find(b);
    if (a == b) return false;
    if (rnk[a] < rnk[b]) swap(a, b);
    par[b] = a;
    if (rnk[a] == rnk[b]) rnk[a]++;
    return true;
}

double px[55], py[55];

int main() {
    int n;
    scanf("%d", &n);
    for (int i = 0; i < n; i++)
        scanf("%lf %lf", &px[i], &py[i]);

    // Build all edges
    vector<Edge> edges;
    for (int i = 0; i < n; i++)
        for (int j = i + 1; j < n; j++) {
            double dx = px[i] - px[j], dy = py[i] - py[j];
            edges.push_back({sqrt(dx*dx + dy*dy), i, j});
        }

    sort(edges.begin(), edges.end());

    // Kruskal's MST
    for (int i = 0; i < n; i++) { par[i] = i; rnk[i] = 0; }

    double totalLen = 0;
    int edgesUsed = 0;
    vector<pair<int,int>> mstEdges;

    for (auto& e : edges) {
        if (unite(e.u, e.v)) {
            totalLen += e.w;
            mstEdges.push_back({e.u, e.v});
            if (++edgesUsed == n - 1) break;
        }
    }

    // Output MST edges and total length
    for (auto& [u, v] : mstEdges)
        printf("%d %d\n", u + 1, v + 1);
    printf("Total length: %.2f\n", totalLen);

    return 0;
}
\end{lstlisting}

\section{Complexity Analysis}

\begin{itemize}
  \item \textbf{Time complexity:} $O(n^2 \log n)$. We generate $\binom{n}{2} = O(n^2)$
        edges and sort them. Kruskal's algorithm with union-by-rank and path compression
        processes each edge in amortized $O(\alpha(n))$ time, where $\alpha$ is the inverse
        Ackermann function. The sorting step dominates.
  \item \textbf{Space complexity:} $O(n^2)$ for storing all edges.
\end{itemize}

\end{document}
