\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red!70!black},
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=8pt,
  frame=single,
  breaklines=true,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 1997 -- Sentiment (Expression Evaluation)}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

Given a logical expression containing variables (single lowercase letters), the operators
AND (\texttt{\&}), OR (\texttt|}), and NOT (\texttt{!}), with parentheses for grouping,
evaluate the expression for given variable assignments.

Operator precedence (highest to lowest): NOT, AND, OR.
Parentheses override precedence.

\medskip
\noindent\textbf{Constraints:} Expression length up to 10{,}000 characters;
variables are single lowercase letters.

\section{Solution Approach}

\subsection{Recursive Descent Parser}

We implement a recursive descent parser that directly mirrors the grammar:

\begin{verbatim}
  expr     -> and_expr ('|' and_expr)*
  and_expr -> not_expr ('&' not_expr)*
  not_expr -> '!' not_expr | atom
  atom     -> variable | '(' expr ')'
\end{verbatim}

Each grammar rule becomes a function that consumes input characters and returns the
boolean value of the corresponding sub-expression. The precedence is naturally encoded:
\texttt{parseExpr} handles OR (lowest precedence), \texttt{parseAnd} handles AND,
and \texttt{parseNot} handles NOT (highest precedence).

\section{C++ Solution}

\begin{lstlisting}
#include <cstdio>
#include <cstring>
using namespace std;

char expr_str[10005];
int pos, len;
int vars[26]; // variable values for 'a'-'z'

int parseExpr();
int parseAnd();
int parseNot();
int parseAtom();

// expr -> and_expr ('|' and_expr)*
int parseExpr() {
    int val = parseAnd();
    while (pos < len && expr_str[pos] == '|') {
        pos++;
        val = val | parseAnd();
    }
    return val;
}

// and_expr -> not_expr ('&' not_expr)*
int parseAnd() {
    int val = parseNot();
    while (pos < len && expr_str[pos] == '&') {
        pos++;
        val = val & parseNot();
    }
    return val;
}

// not_expr -> '!' not_expr | atom
int parseNot() {
    if (pos < len && expr_str[pos] == '!') {
        pos++;
        return !parseNot();
    }
    return parseAtom();
}

// atom -> variable | '(' expr ')'
int parseAtom() {
    if (expr_str[pos] == '(') {
        pos++; // skip '('
        int val = parseExpr();
        pos++; // skip ')'
        return val;
    }
    int idx = expr_str[pos] - 'a';
    pos++;
    return vars[idx];
}

void stripSpaces(char* s) {
    int j = 0;
    for (int i = 0; s[i]; i++)
        if (s[i] != ' ')
            s[j++] = s[i];
    s[j] = '\0';
}

int main() {
    fgets(expr_str, sizeof(expr_str), stdin);
    // Remove trailing newline
    len = strlen(expr_str);
    if (len > 0 && expr_str[len-1] == '\n')
        expr_str[--len] = '\0';
    stripSpaces(expr_str);
    len = strlen(expr_str);

    // Read variable assignments
    int numVars;
    scanf("%d", &numVars);
    for (int i = 0; i < numVars; i++) {
        char c;
        int v;
        scanf(" %c %d", &c, &v);
        vars[c - 'a'] = v;
    }

    pos = 0;
    printf("%d\n", parseExpr());

    return 0;
}
\end{lstlisting}

\section{Correctness}

The parser processes each character exactly once, advancing the global position
\texttt{pos} monotonically. The grammar is unambiguous (left-to-right evaluation
within each precedence level), so the parser correctly evaluates the expression.

Each grammar production handles exactly the operators at its precedence level and
delegates higher-precedence constructs to the next function.

\section{Complexity Analysis}

\begin{itemize}
  \item \textbf{Time complexity:} $O(n)$ where $n$ is the expression length.
        Each character is visited at most once.
  \item \textbf{Space complexity:} $O(n)$ for the recursion stack (worst case:
        deeply nested parentheses or chained NOT operators). $O(1)$ additional space
        for the variable table.
\end{itemize}

\end{document}
