\section*{Overview}

Mo's algorithm is an offline range-query technique for static arrays. Instead of answering queries in the given order,
it reorders them so that the active interval changes only a little between consecutive queries.

The entire method is built around one idea: if I can maintain one interval with cheap \texttt{add} and
\texttt{remove} operations, then smart offline ordering may be enough to answer all queries quickly.

\section*{When to Use It}

Use it when:

\begin{itemize}
  \item the array is static or nearly static,
  \item the statistic is hard for Fenwick trees or segment trees but easy to update when extending/shrinking an
  interval,
  \item offline processing is allowed,
  \item each \texttt{add}/\texttt{remove} is cheap.
\end{itemize}

Typical fits include distinct-value queries, frequency-score queries, and some xor or mode-like statistics.

\section*{Core Idea}

Split indices into blocks of size about \(\sqrt{n}\). Sort queries primarily by the block of the left endpoint and
secondarily by the right endpoint.

Then keep one active interval \([curL, curR]\). For each next query, move the endpoints until they match the desired
interval, calling only:

\begin{itemize}
  \item \texttt{add(index)},
  \item \texttt{remove(index)}.
\end{itemize}

The sorted order is chosen so that the total pointer movement stays manageable.

\section*{Key Insight}

Mo's algorithm does not invent a new clever query formula. It changes the \emph{order} of work so that one maintained
data structure can be reused across many queries.

That makes the add/remove invariant the real center of the method:
\textbf{after each endpoint movement, the maintained statistic must equal the answer on the current active interval.}

\section*{Operations / Main Technique}

The standard recipe is:

\begin{itemize}
  \item choose a block size near \(\sqrt{n}\),
  \item sort queries by Mo order,
  \item maintain the current interval and one answer object,
  \item process sorted queries by endpoint moves.
\end{itemize}

\paragraph{Worked invariant.}
For distinct-value queries, the maintained state is just a frequency table plus one counter \(\texttt{distinct}\). When
adding a value, increment its frequency and raise \(\texttt{distinct}\) if the frequency became \(1\). When removing,
lower \(\texttt{distinct}\) if the frequency became \(0\).

\section*{Correctness Intuition}

Mo's ordering changes only the sequence of queries, not the meaning of any query. At every step, the active interval is
transformed into the next query interval through a sequence of valid add/remove operations.

If the maintained statistic is correct after every single add/remove, then it is correct after the whole transition,
which means it is correct for the new query interval.

\section*{Complexity Analysis}

With block size about \(\sqrt{n}\) and \(O(1)\) add/remove:

\begin{itemize}
  \item sorting: \(O(q \log q)\),
  \item interval movement: about \(O((n + q)\sqrt{n})\),
  \item memory: problem-dependent, usually \(O(n)\) or \(O(\text{value range})\).
\end{itemize}

The constants matter. Mo's algorithm is powerful, but it is not the first structure I try when an online log-time
solution already fits naturally.

\section*{Implementation}

The reference code solves the classic distinct-values query problem. It assumes the values are already coordinate
compressed, which keeps the frequency array compact.

It also uses the alternating right-endpoint order inside each left block. That is a common practical optimization for
reducing pointer movement.

\section*{Common Pitfalls}

\begin{itemize}
  \item Forgetting that the method is offline.
  \item Not coordinate-compressing large values before building the frequency table.
  \item Mixing inclusive intervals with half-open intervals in the pointer moves.
  \item Writing an add/remove pair that is not perfectly symmetric.
  \item Choosing Mo's algorithm even though the problem really wanted a simpler online structure.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Mo with updates, where time becomes a third dimension,
  \item tree Mo after an Euler-tour reduction,
  \item Hilbert-order variants for better constants,
  \item block-size tuning depending on the cost of add/remove.
\end{itemize}

The update and tree variants are powerful, but they are also much easier to get wrong than the plain array version.

\section*{Practice Problems}

\begin{itemize}
  \item Distinct values on subarrays.
  \item Frequency-score queries where add/remove are local.
  \item Offline range problems that resist prefix sums and standard trees.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/data_structures/sqrt_decomposition.html}{cp-algorithms: Sqrt decomposition and Mo's algorithm}.
  \item \href{https://usaco.guide/plat/sqrt}{USACO Guide: Square root decomposition and Mo's algorithm}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
