\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\title{IOI 2020 -- Comparing Plants}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Summary}
Plants are arranged on a circle. The value $r[i]$ describes the local ranking constraint for plant $i$ inside the next $k$ positions. For each query $(x,y)$, we must decide whether every valid height assignment forces $x$ to be taller than $y$, forces $y$ to be taller than $x$, or leaves them incomparable.

\section{Solution}

\subsection{Canonical Permutation}
The official solution first reconstructs one canonical permutation of the plants that is consistent with the array $r$. This is done with:
\begin{itemize}
  \item a segment tree that tracks which positions are currently feasible to place next;
  \item a cyclic ordered set that maintains the gaps between active positions.
\end{itemize}
The result is an ordering $V$ and its inverse position array.

\subsection{Forced Comparisons as Jumps}
Once the canonical order is known, the key observation is that a forced comparison can be propagated clockwise or counterclockwise by jumping to the next position whose canonical rank is strictly larger inside a window of length $k$.

For every start position we compute:
\begin{itemize}
  \item the next forced clockwise jump;
  \item the next forced counterclockwise jump.
\end{itemize}
These jumps are found with Fenwick trees over the canonical ranks. Binary lifting then lets us follow many jumps in $O(\log n)$ time.

\subsection{Answering Queries}
To check whether $a$ is definitely taller than $b$, test whether there is a valid jump chain from $a$ to $b$ either clockwise or counterclockwise. If yes, answer $1$. Otherwise test the reverse direction and answer $-1$ if it holds. If neither direction is forced, answer $0$.

\section{Complexity}
\begin{itemize}
  \item Canonical reconstruction: $O(n \log n)$.
  \item Jump preprocessing: $O(n \log n)$.
  \item Each comparison query: $O(\log n)$.
\end{itemize}

\end{document}
