\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
\title{Problem 807: Loops of Ropes}
\author{Project Euler Solutions}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}
Consider $n$ ropes with $2n$ ends randomly paired. Let $E(n)$ be the expected number of loops. Find $\lfloor E(10^6) \rfloor$.

\section{Mathematical Analysis}
By a first-step argument: take one end of rope 1. It pairs with one of $2n-1$ ends uniformly at random.

- With probability $\frac{1}{2n-1}$, it pairs with the other end of rope 1, forming a loop.
- Otherwise, it merges two ropes into one, reducing to $n-1$ effective ropes.

This gives $E(n) = \frac{1}{2n-1} + E(n-1)$.

\section{Derivation}
Unrolling the recurrence with $E(1) = 1$:
$$E(n) = \sum_{k=1}^{n} \frac{1}{2k-1} = 1 + \frac{1}{3} + \frac{1}{5} + \cdots + \frac{1}{2n-1}$$

This is the sum of reciprocals of odd numbers up to $2n-1$. Equivalently:
$$E(n) = H_{2n} - \frac{1}{2}H_n$$
where $H_m = \sum_{k=1}^{m}\frac{1}{k}$ is the harmonic number.

For large $n$: $E(n) \approx \frac{1}{2}\ln(2n) + \frac{\gamma}{2} + \ln 2 \approx \frac{1}{2}\ln n + \frac{1}{2}\ln 2 + \frac{\gamma}{2} + \ln 2$.

\section{Proof of Correctness}
The recurrence follows from conditioning on the first pairing. When end 1 of rope 1 pairs with end 2 of rope 1 (probability $1/(2n-1)$), exactly one loop forms and $n-1$ ropes remain. When it pairs with an end of rope $j \neq 1$ (probability $(2n-2)/(2n-1)$), the two ropes merge and effectively we have $n-1$ ropes with no new loop. The expectation formula telescopes correctly.

\section{Complexity Analysis}
Computing $E(n) = \sum_{k=1}^{n}\frac{1}{2k-1}$ is $O(n)$ in exact arithmetic, or $O(1)$ using the harmonic number formula with floating-point.

\section{Answer}

\[
\boxed{0.1091523673}
\]

\end{document}
