¥documentclass[11pt]{article}
¥usepackage{amsmath,amssymb,amsthm}
¥usepackage{algorithm2e}
¥usepackage[margin=1in]{geometry}

¥newtheorem{theorem}{Theorem}
¥newtheorem{lemma}[theorem]{Lemma}

¥title{Problem 671: Colouring a Loop}
¥author{Project Euler Solutions}
¥date{}

¥begin{document}
¥maketitle

¥section{Problem Statement}

Extend the tiling of Problem‾670 to a $2 ¥times n$ ¥textbf{loop} (cylinder). Let $F_k(n)$ count the number of valid colourings of a $2 ¥times n$ cylindrical grid using at most $k$ colours, where adjacent cells must have different colours. Given:
¥begin{itemize}
  ¥item $F_4(3) = 104$,
  ¥item $F_5(7) = 3¥,327¥,300$,
  ¥item $F_6(101) ¥equiv 75¥,309¥,980 ¥pmod{1¥,000¥,004¥,321}$.
¥end{itemize}
Find $F_{10}(10¥,004¥,003¥,002¥,001) ¥bmod 1¥,000¥,004¥,321$.

¥section{Mathematical Foundation}

Let $¥mathcal{S}$ denote the set of valid column states for a $2 ¥times 1$ column with $k$ colours, and let $T$ be the $|¥mathcal{S}| ¥times |¥mathcal{S}|$ transfer matrix where $T_{ij} = 1$ if column state‾$j$ can follow column state‾$i$.

¥begin{theorem}[Trace Formula for Cylindrical Tilings]
The number of valid $k$-colourings of a $2 ¥times n$ cylinder is
¥[
  F_k(n) = ¥operatorname{tr}(T^n) = ¥sum_{i=1}^{s} ¥lambda_i^n,
¥]
where $¥lambda_1, ¥ldots, ¥lambda_s$ are the eigenvalues of‾$T$ counted with algebraic multiplicity, and $s = |¥mathcal{S}|$.
¥end{theorem}

¥begin{proof}
A valid colouring of the cylinder is a closed walk of length‾$n$ in the state graph: a sequence of states $q_0, q_1, ¥ldots, q_{n-1}$ with $q_0 = q_{n-1}$ such that each consecutive pair is compatible. The number of closed walks of length‾$n$ starting and ending at state‾$i$ is $(T^n)_{ii}$. Summing over all starting states yields $¥sum_i (T^n)_{ii} = ¥operatorname{tr}(T^n)$. By diagonalisation (or Jordan normal form), $¥operatorname{tr}(T^n) = ¥sum_i ¥lambda_i^n$.
¥end{proof}

¥begin{theorem}[Cayley--Hamilton Recurrence]
Let $¥chi_T(¥lambda) = ¥lambda^s - c_1 ¥lambda^{s-1} - ¥cdots - c_s$ be the characteristic polynomial of‾$T$. Then the sequence $a_n = ¥operatorname{tr}(T^n)$ satisfies the linear recurrence
¥[
  a_n = c_1¥, a_{n-1} + c_2¥, a_{n-2} + ¥cdots + c_s¥, a_{n-s} ¥quad ¥text{for all } n ¥ge s.
¥]
¥end{theorem}

¥begin{proof}
By the Cayley--Hamilton theorem, $T^s = c_1 T^{s-1} + ¥cdots + c_s I$. Taking the trace of $T^n = T^{n-s} ¥cdot T^s$ and applying linearity of trace:
¥[
  ¥operatorname{tr}(T^n) = c_1 ¥operatorname{tr}(T^{n-1}) + ¥cdots + c_s ¥operatorname{tr}(T^{n-s}). ¥qedhere
¥]
¥end{proof}

¥begin{lemma}[Modular Exponentiation of Recurrence]
Given a linear recurrence of order‾$s$ over $¥mathbb{Z}/p¥mathbb{Z}$, the $n$-th term can be computed in $O(s^2 ¥log n)$ arithmetic operations using Kitamasa's method.
¥end{lemma}

¥begin{proof}
Represent the shift operator as multiplication by‾$x$ in the quotient ring $(¥mathbb{Z}/p¥mathbb{Z})[x]/(¥chi_T(x))$. Computing $x^n ¥bmod ¥chi_T(x)$ by repeated squaring requires $O(¥log n)$ polynomial multiplications, each costing $O(s^2)$. Reading off‾$a_n$ from the resulting polynomial and the initial values is $O(s)$.
¥end{proof}

¥section{Algorithm}

¥begin{enumerate}
  ¥item Build the transfer matrix‾$T$ of size $s ¥times s$ for $k = 10$ colours.
  ¥item Compute the characteristic polynomial $¥chi_T$ modulo $p = 1¥,000¥,004¥,321$.
  ¥item Compute initial values $a_0, ¥ldots, a_{s-1}$ via iterated matrix--vector multiplication.
  ¥item Use Kitamasa's method: compute $x^n ¥bmod ¥chi_T(x)$ in $¥mathbb{F}_p[x]$ by repeated squaring.
  ¥item Recover $F_k(n) = ¥sum_j [¥text{coeff}_j] ¥cdot a_j ¥bmod p$.
¥end{enumerate}

¥section{Complexity Analysis}

¥begin{itemize}
  ¥item ¥textbf{Time:} $O(s^2 k^2)$ to build‾$T$, $O(s^3)$ for the characteristic polynomial, $O(s^2 ¥log n)$ for Kitamasa's method. Total: $O(k^6 + k^4 ¥log n)$.
  ¥item ¥textbf{Space:} $O(s^2) = O(k^4)$ for the transfer matrix.
¥end{itemize}

¥section{Answer}

¥[
¥boxed{946106780}
¥]

¥end{document}
