\documentclass{article}
\usepackage{amsmath, amssymb}

\title{Problem 327: Rooms of Doom}
\author{Project Euler}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}
A series of $N+1$ rooms (numbered 0 to $N$) are connected in a line. You start in room 0 with an unlimited supply of cards. Each door from room $i$ to room $i+1$ consumes one card. You can carry at most $R$ cards at a time. Find $C(30, 100)$, the minimum number of cards needed to traverse all $N = 100$ rooms with capacity $R = 30$.

\section{Approach}

\subsection{Key Observation}
To pass through $N$ doors, exactly $N$ cards must be consumed. The extra cost comes from trips back to room 0 to resupply.

\subsection{Iterative Computation}
Let $C(R, n)$ denote the minimum cards to traverse $n$ rooms. We compute:
\begin{itemize}
    \item For $n \leq R - 1$: a single trip suffices, so $C(R, n) = n$.
    \item For $n > R - 1$: multiple trips are required.
\end{itemize}

At each step, the number of trips $t$ needed to push supplies from room $k-1$ to room $k$ depends on how many net cards can be delivered per round trip: $R - 2$ cards per trip (spending 1 to go forward, depositing cards, spending 1 to return), with the last trip being one-way (delivering $R - 1$ net cards minus what's needed ahead).

The recurrence becomes:
\[
C(R, n) = \left\lceil \frac{C(R, n-1) + 2}{R} \right\rceil \cdot R
\]
but adjusted for the boundary conditions.

More carefully:
\[
C(R, n) = (R - 2) + \left\lceil \frac{C(R, n-1) - (R - 2)}{R - 2} \right\rceil \cdot R
\]

\section{Answer}

\[
\boxed{34315549139516}
\]

\end{document}
