\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage[margin=2.5cm]{geometry}
\usepackage{algorithm}
\usepackage{algpseudocode}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\title{Project Euler Problem 405:\\A Rectangular Tiling}
\author{Solution Write-up}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

We tile a rectangle whose length is twice its width using a recursive substitution
rule.  Define:
\begin{itemize}[nosep]
  \item $T(0)$: the tiling consisting of a single $2\times 1$ rectangle.
  \item $T(n)$ for $n>0$: obtained from $T(n-1)$ by replacing every tile at
        double scale with tiles of the orthogonal orientation (a brick-like pattern).
\end{itemize}

Let $f(n)$ denote the number of interior points where exactly four tiles meet
in~$T(n)$.

\medskip\noindent
\textbf{Given:} $f(1)=0$,\; $f(4)=82$,\; $f(10^9)\bmod 17^7 = 126\,897\,180$.

\medskip\noindent
\textbf{Find:} $f(10^k)$ for $k=10^{18}$, modulo $17^7 = 410\,338\,673$.

\section{Substitution Rule}

Each $2{:}1$ tile is scaled to $4{:}2$ and subdivided as follows:

\begin{itemize}[nosep]
  \item A \emph{horizontal} $2\times 1$ tile becomes a $4\times 2$ block
        containing two vertical $1\times 2$ tiles on the sides and two
        horizontal $2\times 1$ tiles stacked in the centre.
  \item A \emph{vertical} $1\times 2$ tile becomes a $2\times 4$ block
        containing two horizontal $2\times 1$ tiles at top and bottom and
        two vertical $1\times 2$ tiles stacked in the centre.
\end{itemize}

Each substitution step produces $4^n$ tiles in a grid of size
$2^{n+1}\times 2^n$.  Because the rule alternates orientations, no
four-corner point is created within a single substituted block; all
four-corners arise at boundaries between blocks.

\section{Simulation of Small Values}

By explicit construction and counting:

\begin{center}
\begin{tabular}{crr}
\toprule
$n$ & Tiles & $f(n)$ \\
\midrule
0 & 1     & 0 \\
1 & 4     & 0 \\
2 & 16    & 2 \\
3 & 64    & 16 \\
4 & 256   & 82 \\
5 & 1024  & 368 \\
6 & 4096  & 1554 \\
7 & 16384 & 6384 \\
\bottomrule
\end{tabular}
\end{center}

\section{Linear Recurrence}

\begin{theorem}
The sequence $f(n)$ satisfies
\[
  f(n) \;=\; 5\,f(n-1) \;-\; 2\,f(n-2) \;-\; 8\,f(n-3) \;+\; 6,
  \qquad n\ge 4,
\]
with initial values $f(1)=0$, $f(2)=2$, $f(3)=16$.
\end{theorem}

\begin{proof}
Verified computationally for $n=4,\dots,7$ against the simulation.
The characteristic polynomial of the homogeneous part is
\[
  x^3 - 5x^2 + 2x + 8 = (x-4)(x-2)(x+1),
\]
so the dominant growth rate is~$4^n$, consistent with the quadrupling of
the grid area at each step.
\end{proof}

\section{Closed-Form Solution}

The general solution is $f(n) = A\cdot 4^n + B\cdot 2^n + C\cdot(-1)^n + D$
where $D=1$ is the particular solution for the constant forcing term.
Solving the initial conditions:

\[
  A = \tfrac{2}{5},\quad B = -\tfrac{4}{3},\quad C = -\tfrac{1}{15}.
\]

Hence
\[
  \boxed{f(n) \;=\; \frac{6\cdot 4^n \;-\; 20\cdot 2^n \;-\; (-1)^n \;+\; 15}{15}.}
\]

\begin{lemma}
For all $n\ge 1$, the numerator $6\cdot 4^n - 20\cdot 2^n - (-1)^n + 15$ is
divisible by~$15$.
\end{lemma}

\begin{proof}
Modulo~$3$: $6\cdot 4^n \equiv 0$, $20\cdot 2^n \equiv 2\cdot 2^n$,
$(-1)^n + 15 \equiv (-1)^n$.  So the numerator $\equiv -2\cdot 2^n - (-1)^n
\equiv 0\pmod{3}$ (check $n$ odd and even separately).
Similarly modulo~$5$.
\end{proof}

\section{Matrix Exponentiation}

For modular computation we cast the recurrence in matrix form:
\[
  \begin{pmatrix} f(n{+}1) \\ f(n) \\ f(n{-}1) \\ 1 \end{pmatrix}
  =
  \underbrace{\begin{pmatrix}
    5 & -2 & -8 & 6 \\
    1 &  0 &  0 & 0 \\
    0 &  1 &  0 & 0 \\
    0 &  0 &  0 & 1
  \end{pmatrix}}_{M}
  \begin{pmatrix} f(n) \\ f(n{-}1) \\ f(n{-}2) \\ 1 \end{pmatrix}.
\]
Then $f(n)$ is the first component of $M^{n-3}\,\mathbf{v}$ where
$\mathbf{v} = (16,\,2,\,0,\,1)^T$.

\section{Computing \texorpdfstring{$f\bigl(10^{(10^{18})}\bigr) \bmod 17^7$}{f(10\^(10\^18)) mod 17\^7}}

\subsection{Order of $M$ modulo $17^7$}

\begin{proposition}
The multiplicative order of $M$ modulo $17^e$ is
\[
  \mathrm{ord}_{17^e}(M) =
  \begin{cases}
    8 & \text{if } e=1, \\
    8\cdot 17^{e-1} & \text{if } e\ge 2.
  \end{cases}
\]
In particular, $\mathrm{ord}_{17^7}(M) = 8\cdot 17^6 = 193\,100\,552$.
\end{proposition}

\subsection{Exponent Reduction}

We need $N - 3 \pmod{\mathrm{ord}}$ where $N = 10^{10^{18}}$ and
$\mathrm{ord} = 2^3 \cdot 17^6$.

\paragraph{Modulo $8$:}
$10^k \equiv 0\pmod{8}$ for $k\ge 3$.  Since $10^{18}\ge 3$, we have
$N\equiv 0\pmod{8}$.

\paragraph{Modulo $17^6$:}
By Euler's theorem, $10^{\varphi(17^6)}\equiv 1\pmod{17^6}$ with
$\varphi(17^6)=16\cdot 17^5 = 22\,717\,712$.  Thus
\[
  10^{10^{18}} \bmod 17^6 \;=\; 10^{\,10^{18}\!\bmod\,22\,717\,712}\!\bmod 17^6.
\]
Compute $10^{18}\bmod 22\,717\,712 = 16\,672\,816$, then
$10^{16\,672\,816}\bmod 17^6 = 7\,544\,618$.

\paragraph{CRT:}
Combining: $N\bmod\,\mathrm{ord} = 152\,370\,032$.

The matrix exponent is $(152\,370\,032 - 3)\bmod\,\mathrm{ord} = 152\,370\,029$.

\subsection{Closed-Form Evaluation}

Alternatively, using the closed form directly:
\begin{align*}
  f(N) &\equiv \bigl(6\cdot 4^N - 20\cdot 2^N - (-1)^N + 15\bigr)
         \cdot 15^{-1} \pmod{17^7}.
\end{align*}
Since $N=10^{10^{18}}$ is even, $(-1)^N=1$.  The powers $2^N$ and $4^N=2^{2N}$
are reduced modulo $17^7$ using $\varphi(17^7) = 386\,201\,104$.

Both the matrix method and the closed-form method yield the same answer.

\section{Answer}

\[
\boxed{237696125}
\]
\section{Complexity}

\begin{itemize}[nosep]
  \item \textbf{Closed-form:} $O(\log N \cdot \log m)$ for modular exponentiation,
        which after exponent reduction is $O(\log^2 m)$.
  \item \textbf{Matrix method:} $O(4^3 \log\mathrm{ord})$ $= O(64\log m)$.
  \item Both run in well under one second.
\end{itemize}

\end{document}
