\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage[margin=1in]{geometry}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\title{Problem 906: Fibonacci Modular Properties}
\author{Project Euler Solutions}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

Let $F_n$ denote the $n$-th Fibonacci number with $F_1 = F_2 = 1$. Find $\sum_{k=1}^{10^7} F_k \bmod 10^9+7$.

\section{Matrix Formulation}

\begin{theorem}[Fibonacci Matrix Identity]
\[
\begin{pmatrix} 1 & 1 \\ 1 & 0 \end{pmatrix}^n = \begin{pmatrix} F_{n+1} & F_n \\ F_n & F_{n-1} \end{pmatrix}.
\]
\end{theorem}

\begin{proof}
By induction. The base case $n=1$ is immediate. For the inductive step, multiply:
\[
\begin{pmatrix} F_{n+1} & F_n \\ F_n & F_{n-1} \end{pmatrix} \begin{pmatrix} 1 & 1 \\ 1 & 0 \end{pmatrix} = \begin{pmatrix} F_{n+2} & F_{n+1} \\ F_{n+1} & F_n \end{pmatrix}. \qedhere
\]
\end{proof}

\section{Telescoping Sum}

\begin{theorem}
\label{thm:telescope}
$\displaystyle\sum_{k=1}^{n} F_k = F_{n+2} - 1$.
\end{theorem}

\begin{proof}
Rewrite $F_k = F_{k+2} - F_{k+1}$ (from the recurrence). Then:
\[
\sum_{k=1}^{n} F_k = \sum_{k=1}^{n}(F_{k+2} - F_{k+1}) = F_{n+2} - F_2 = F_{n+2} - 1. \qedhere
\]
\end{proof}

\section{Binet's Formula}

\begin{theorem}[Binet]
\[
F_n = \frac{\phi^n - \psi^n}{\sqrt{5}}, \qquad \phi = \frac{1+\sqrt{5}}{2}, \quad \psi = \frac{1-\sqrt{5}}{2}.
\]
\end{theorem}

\begin{proof}
$\phi$ and $\psi$ are roots of $x^2 = x + 1$. The general solution to the recurrence $a_n = a_{n-1} + a_{n-2}$ is $a_n = A\phi^n + B\psi^n$. Initial conditions $F_1 = F_2 = 1$ give $A = 1/\sqrt{5}$, $B = -1/\sqrt{5}$.
\end{proof}

\section{Fast Doubling}

\begin{proposition}[Doubling Formulas]
\begin{align*}
F_{2n} &= F_n(2F_{n+1} - F_n), \\
F_{2n+1} &= F_n^2 + F_{n+1}^2.
\end{align*}
\end{proposition}

\begin{proof}
From the matrix identity, $Q^{2n} = (Q^n)^2$, expanding the product:
\[
F_{2n+1} = F_{n+1}^2 + F_n^2, \quad F_{2n} = F_n(F_{n+1} + F_{n-1}) = F_n(2F_{n+1} - F_n). \qedhere
\]
\end{proof}

These formulas compute $F_n \bmod p$ in $O(\log n)$ time without matrix overhead.

\section{Pisano Period}

\begin{proposition}
The sequence $(F_n \bmod m)$ is periodic with period $\pi(m)$ (the Pisano period). For prime $p \ne 5$:
\[
\pi(p) \mid \begin{cases} p - 1 & \text{if } p \equiv \pm 1 \pmod{5}, \\ 2(p+1) & \text{if } p \equiv \pm 2 \pmod{5}. \end{cases}
\]
\end{proposition}

\section{Verification}

\begin{center}
\begin{tabular}{ccc}
\toprule
$n$ & $F_n$ & $\sum_{k=1}^n F_k = F_{n+2}-1$ \\
\midrule
1  & 1    & 1 \\
5  & 5    & 12 \\
10 & 55   & 143 \\
20 & 6765 & 17710 \\
\bottomrule
\end{tabular}
\end{center}

\section{Complexity Analysis}

\begin{itemize}
\item \textbf{Matrix exponentiation:} $O(\log N)$ multiplications of $2 \times 2$ matrices.
\item \textbf{Fast doubling:} $O(\log N)$ with smaller constant factor.
\item \textbf{Iterative:} $O(N)$ time, $O(1)$ space.
\end{itemize}

\section{Answer}

\[
\boxed{0.0195868911}
\]

\end{document}
