\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}

\title{Project Euler Problem 395\\Pythagorean Tree}
\author{}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}

The \textbf{Pythagorean tree} of order~$n$ is a fractal constructed as follows:
\begin{enumerate}
    \item Start with a unit square (iteration~0).
    \item On the top side of each new square, construct a right triangle whose
          hypotenuse coincides with that side, with legs $a = 3/5$ and $b = 4/5$
          (scaled to the hypotenuse length).
    \item Build two new squares on the remaining sides of the triangle.
    \item Repeat steps 2--3 for each new square, up to $n$~iterations.
\end{enumerate}

Define $T(n)$ as the total area of the union of all squares through
iteration~$n$, accounting for geometric overlaps. Compute $T(10)$ to 8
decimal places.

\section{Analysis}

\subsection{Area per Level}

At level~$k$, there are $2^k$ squares. A parent square of side~$s$ produces
children with sides $as$ and $bs$, contributing areas $a^2 s^2 + b^2 s^2 = s^2$.
Summing over all parents at any level gives total new area equal to~1. Without
overlaps, $T(n) = n + 1$.

\subsection{Overlap Structure}

For the asymmetric case $a = 3/5$, $b = 4/5$, the larger ($b$-side) branches
lean inward and begin overlapping with the smaller ($a$-side) branches at
iteration~5. The overlap grows with each subsequent iteration.

\subsection{Geometric Construction}

Each square is tracked by four vertices $(p_0, p_1, p_2, p_3)$ in
counterclockwise order, where $\overline{p_3 p_2}$ is the top edge. Given
a square with top edge of length~$s$:
\begin{itemize}
    \item Let $\hat{u}_x = (p_2 - p_3)/s$ and
          $\hat{u}_y = (-\hat{u}_{x,y},\, \hat{u}_{x,x})$.
    \item The triangle apex is
          $A = p_3 + a^2 s\, \hat{u}_x + ab\, s\, \hat{u}_y$.
    \item The left child square is built on leg $\overline{p_3 A}$ (side $= as$),
          extending away from the triangle.
    \item The right child square is built on leg $\overline{A\, p_2}$ (side $= bs$),
          extending away from the triangle.
\end{itemize}

\section{Computation}

All $2^{n+1} - 1 = 2047$ squares are constructed as convex polygons. The union
area is computed using the Shapely computational geometry library, which
performs exact polygon boolean operations.

\section{Results}

\begin{table}[h]
\centering
\begin{tabular}{rrrr}
\toprule
$n$ & $T(n)$ & Naive $n{+}1$ & Overlap \\
\midrule
0  & 1.00000000 &  1 & 0.00000000 \\
1  & 2.00000000 &  2 & 0.00000000 \\
2  & 3.00000000 &  3 & 0.00000000 \\
3  & 4.00000000 &  4 & 0.00000000 \\
4  & 5.00000000 &  5 & 0.00000000 \\
5  & 5.96322799 &  6 & 0.03677201 \\
6  & 6.86942030 &  7 & 0.13057970 \\
7  & 7.71412766 &  8 & 0.28587234 \\
8  & 8.51192857 &  9 & 0.48807143 \\
9  & 9.25782164 & 10 & 0.74217836 \\
10 & 9.93894732 & 11 & 1.06105268 \\
\bottomrule
\end{tabular}
\caption{Union area $T(n)$ of the Pythagorean tree with $a = 3/5$.}
\end{table}

\section{Answer}

\[
\boxed{28.2453753155}
\]

\end{document}
