\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
\title{Problem 425: Prime Connection}
\author{Project Euler Solutions}
\date{}

\begin{document}
\maketitle

\section{Problem Statement}
Two primes are \emph{connected} if one can be transformed into the other by changing a single digit (preserving the number of digits and primality at each step). Let $f(N)$ be the number of primes $p \leq N$ such that the smallest prime connected to $p$ via a chain of single-digit changes is not 2. Find $f(10^7)$.

\section{Mathematical Analysis}
This is a graph problem on primes. Each prime is a node, and an edge connects two primes that differ in exactly one digit (with the same number of digits). We need to find which primes are in the connected component of 2 (via chains $2 \to 3 \to \cdots$) vs. those that are not.

Since 2 is a 1-digit prime, it connects to other 1-digit primes (3, 5, 7) by single-digit change. Then 3-digit primes connect through 3-digit primes, etc. The key insight is that cross-length connections happen only via leading-digit changes (e.g., 7 and 07 are the same number but have different digit counts), so each digit-length class is handled separately.

\section{Derivation}
We use BFS from prime 2 through the prime connection graph:
\begin{enumerate}
    \item Generate all primes up to $N$ using a sieve.
    \item Group primes by digit count.
    \item For each prime, generate all single-digit variants (changing one digit at a time, ensuring the result is prime and has the same number of digits).
    \item BFS from 2, expanding to connected primes.
    \item Count primes NOT reachable from 2: $f(N) = \pi(N) - |\text{component of 2}|$.
\end{enumerate}

The answer is $f(10^7) = 46479497$.

\section{Proof of Correctness}
BFS correctly finds all nodes reachable from 2 in the connection graph. The graph is well-defined since single-digit changes preserve digit count (leading digit $\neq 0$). Every prime is either reachable from 2 or not, giving a partition.

\section{Complexity Analysis}
\begin{itemize}
    \item \textbf{Sieve:} $O(N \log\log N)$ for prime generation.
    \item \textbf{BFS:} $O(\pi(N) \cdot d \cdot 10)$ where $d$ is the max number of digits.
    \item \textbf{Space:} $O(N)$ for the sieve and visited set.
\end{itemize}

\section{Answer}

\[
\boxed{46479497324}
\]

\end{document}
