\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}

\title{ICPC World Finals 2021\\I. Spider Walk}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Problem Summary}

Charlotte starts on one strand, walks outward, and is forced to cross every existing bridge she meets.
For each starting strand, we must compute the minimum number of \emph{new} bridges needed so that she
eventually ends on strand $s$.

The key point is that a new bridge can be placed anywhere between two consecutive existing bridge
distances, so within one such annulus Charlotte may switch to a neighboring strand whenever she wants,
paying cost $1$ each time.

\section*{Key Observations}

\begin{itemize}[leftmargin=*]
    \item Instead of thinking about geometric strands, think about the \emph{autopilot trajectories} of all
    $n$ starting strands. As radius increases, these trajectories cross whenever an existing bridge swaps
    two adjacent strands.

    \item Each trajectory is a polyline from an inner endpoint to an outer endpoint. Moving along a
    trajectory costs $0$.

    \item Between two consecutive bridge distances, the order of trajectories is fixed. In that annulus,
    Charlotte may add a new bridge between any two adjacent geometric strands, which is exactly the same
    as switching between two adjacent trajectories. Each such switch costs $1$.

    \item Therefore the whole problem becomes a shortest-path problem in a graph:
    \begin{itemize}[leftmargin=*]
        \item vertices represent trajectory segments between consecutive crossings,
        \item directed edges of cost $0$ connect consecutive segments of the same trajectory from inner to
        outer radius,
        \item bidirectional edges of cost $1$ connect adjacent trajectory segments within one annulus.
    \end{itemize}

    \item We never need one vertex per strand per annulus. A new vertex is created only when a trajectory
    itself crosses a bridge or one of its adjacent trajectories changes, so each bridge causes only $O(1)$
    fresh vertices.
\end{itemize}

\section*{Algorithm}

\begin{enumerate}[leftmargin=*]
    \item Sort all existing bridges by increasing distance from the center.

    \item Sweep outward and maintain:
    \begin{itemize}[leftmargin=*]
        \item \texttt{order[pos]}: which starting trajectory currently occupies geometric position
        \texttt{pos},
        \item \texttt{active[label]}: the current graph vertex for the still-open segment of that trajectory.
    \end{itemize}

    \item Initially, each starting strand contributes one graph vertex. Since all trajectories are adjacent in
    the initial annulus, add cost-$1$ edges between consecutive positions on the cycle.

    \item When processing a bridge on strands $t$ and $t+1$ (cyclically):
    \begin{itemize}[leftmargin=*]
        \item first swap the two trajectories in \texttt{order}, because this describes the new annulus just
        outside the bridge;
        \item the only trajectories whose set of adjacent trajectories may have changed are those currently
        occupying positions
        \[
            \text{prev}(t),\ t,\ t+1,\ \text{next}(t+1).
        \]
        For each distinct such position, create a new graph vertex for the trajectory now occupying it and
        connect its previous vertex to this new vertex by a directed cost-$0$ edge;
        \item in the new annulus, every adjacency involving one of these refreshed positions starts a new
        interval, so add cost-$1$ edges for all cyclic adjacent pairs touching them.
    \end{itemize}

    \item After the sweep, the trajectory currently occupying outer strand $s$ is the target. Run $0$-$1$
    BFS on the \emph{reversed} graph from its final graph vertex.

    \item The answer for starting strand $i$ is the distance from this target to the initial graph vertex of
    trajectory $i$.
\end{enumerate}

\section*{Correctness Proof}

We prove that the algorithm returns the correct answer.

\paragraph{Lemma 1.}
Every valid movement of Charlotte corresponds to a path in the constructed graph with the same cost.

\paragraph{Proof.}
Whenever Charlotte follows the web without adding a bridge, she stays on one autopilot trajectory. In
the graph this is exactly moving along cost-$0$ edges between consecutive segments of that trajectory.

If Charlotte adds a bridge in some annulus, she switches between two adjacent geometric strands. Since
the order of trajectories in that annulus is fixed, this is precisely a switch between two adjacent
trajectory segments, represented by one cost-$1$ edge.

Thus every physical walk maps to a graph path whose total cost equals the number of added bridges. \qed

\paragraph{Lemma 2.}
Every path in the constructed graph corresponds to a valid strategy for Charlotte with the same cost.

\paragraph{Proof.}
Cost-$0$ edges connect consecutive pieces of the same trajectory across a crossing, so traversing them
means simply continuing outward and taking the required existing bridge.

Cost-$1$ edges connect two adjacent trajectory segments that coexist in one annulus. Charlotte can place a
new bridge anywhere inside that annulus and switch exactly once between those two trajectories. Therefore
every graph path can be realized physically, with one added bridge for each cost-$1$ edge. \qed

\paragraph{Lemma 3.}
The graph has exactly the transitions needed to describe all possible strategies.

\paragraph{Proof.}
Initially each trajectory has one segment. Every existing bridge splits exactly the two trajectories that
cross there, and may also change the local neighbors of the two immediately adjacent trajectories. Hence
after one bridge, only the four cyclic positions
\[
    \text{prev}(t),\ t,\ t+1,\ \text{next}(t+1)
\]
need fresh segment vertices. So the sweep creates only $O(1)$ new vertices per bridge.

Within a fixed annulus, only adjacent trajectories can be connected by a newly added bridge, so the
graph contains exactly those cost-$1$ edges. A bridge crossing changes only local adjacency, hence only
a constant number of adjacency pairs need to be updated after each event. No other transitions are
created or lost. \qed

\paragraph{Theorem.}
The algorithm outputs the correct answer for every valid input.

\paragraph{Proof.}
By Lemmas 1 and 2, minimum added bridges from a starting strand to outer strand $s$ are exactly equal
to shortest-path distance in the directed constructed graph from the initial segment of that trajectory to
the final segment that exits at strand $s$. By Lemma 3, the graph contains all and only valid transitions.
Running $0$-$1$ BFS on the reversed graph from the target computes these same shortest-path distances to
all starting vertices at once. Therefore each printed distance is correct. \qed

\section*{Complexity Analysis}

Sorting the bridges takes $O(m \log m)$ time.

The graph has $O(n+m)$ vertices and edges:
\begin{itemize}[leftmargin=*]
    \item $O(n+4m)$ trajectory-segment vertices,
    \item $O(n+m)$ zero-cost and unit-cost edges.
\end{itemize}

The sweep is linear after sorting, and $0$-$1$ BFS is also linear in the graph size. Therefore the total
time complexity is
\[
    O(m \log m + n + m),
\]
and memory usage is
\[
    O(n+m).
\]

\section*{Implementation Notes}

\begin{itemize}[leftmargin=*]
    \item Strand $n$ is adjacent to strand $1$, so all position updates are cyclic.
    \item Multiple cost-$1$ edges between the same pair of vertices are harmless; $0$-$1$ BFS handles
    them naturally.
    \item The target is the \emph{trajectory} that ends at strand $s$, not necessarily the trajectory that
    started at strand $s$.
\end{itemize}

\end{document}
