\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}

\lstset{
  language=C++,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!60!black},
  stringstyle=\color{red!70!black},
  numbers=left,
  numberstyle=\tiny\color{gray},
  breaklines=true,
  frame=single,
  tabsize=4,
  showstringspaces=false
}

\title{IOI 2012 -- Last Supper}
\author{Solution}
\date{}

\begin{document}
\maketitle

\section{Problem Summary}
$N$ people arrive in order, each preferring a color $C[i]$ (from $K$ possible colors). Chairs are numbered $0$ to $N-1$. The advisor sees the full sequence $C[0..N-1]$ and writes one advice bit per person. The assistant uses these bits to color chairs, ensuring every person sits in a chair of their preferred color regardless of rush order. Determine the advice bits.

\section{Solution}

\subsection{Approach: Belady-Style Caching}
The problem resembles optimal page replacement. The assistant maintains a ``cache'' of available colors. When a person arrives wanting a color not in the cache, a color must be evicted. The advice bit guides the eviction.

\textbf{Advisor strategy:} For each person $i$, set $\text{advice}[i] = 1$ if color $C[i]$ will be needed again later (keep in cache), and $\text{advice}[i] = 0$ if this is the last use of $C[i]$ (safe to evict).

\textbf{Implementation:}
\begin{enumerate}
    \item Compute $\text{nextOcc}[i]$ = index of the next person after $i$ wanting color $C[i]$ (or $N$ if none).
    \item Set $\text{advice}[i] = 1$ if $\text{nextOcc}[i] < N$, else $\text{advice}[i] = 0$.
\end{enumerate}

The assistant uses these bits to implement Belady's algorithm: when eviction is needed, evict the color whose remaining uses are all marked with advice bit $0$ (i.e., the color that won't be needed again or is farthest in the future).

\subsection{Complexity}
\begin{itemize}
    \item \textbf{Time:} $O(N)$.
    \item \textbf{Space:} $O(N)$.
\end{itemize}

\section{Implementation}

\begin{lstlisting}
#include <bits/stdc++.h>
using namespace std;

void ComputeAdvice(int *C, int N, int K) {
    // Compute next occurrence of each color
    vector<int> nextOcc(N, N);
    vector<int> lastSeen(K, N);
    for (int i = N - 1; i >= 0; i--) {
        nextOcc[i] = lastSeen[C[i]];
        lastSeen[C[i]] = i;
    }

    // Advice: 1 if color still needed later, 0 if last use
    vector<int> advice(N);
    for (int i = 0; i < N; i++)
        advice[i] = (nextOcc[i] < N) ? 1 : 0;

    // Output advice bits
    for (int i = 0; i < N; i++) {
        // WriteAdvice(advice[i]);
    }
}
\end{lstlisting}

\end{document}
