\section*{Overview}

Euler's phi function \(\varphi(n)\) counts how many integers in \([1, n]\) are coprime to \(n\). It shows up in
modular arithmetic, counting reduced fractions, cycle lengths, and several multiplicative-function problems.

\section*{When to Use It}

Use \(\varphi\) when:

\begin{itemize}
  \item the question asks for coprime counts,
  \item Euler's theorem might replace a large exponent,
  \item a multiplicative formula over prime factors is wanted.
\end{itemize}

\section*{Core Idea}

If
\[
n = p_1^{e_1} p_2^{e_2} \cdots p_k^{e_k},
\]
then
\[
\varphi(n) = n \prod_{i=1}^{k}\left(1 - \frac{1}{p_i}\right).
\]

So once the distinct prime divisors are known, computing \(\varphi(n)\) is straightforward.

\section*{Key Insight}

Phi is counting by exclusion. Among the numbers \(1\) through \(n\), the only ones that fail to be coprime to \(n\) are
those divisible by one of its prime factors. The product formula is inclusion-exclusion compressed into multiplicative
form.

\section*{Operations / Main Technique}

\begin{itemize}
  \item factor \(n\),
  \item start from \(\texttt{result = n}\),
  \item for each distinct prime divisor \(p\), apply
  \[
  \texttt{result -= result / p}.
  \]
\end{itemize}

\paragraph{Worked example.}
For \(n = 12 = 2^2 \cdot 3\),
\[
\varphi(12) = 12\left(1 - \frac{1}{2}\right)\left(1 - \frac{1}{3}\right) = 4.
\]
The coprime values are \(1, 5, 7, 11\).

\section*{Correctness Intuition}

Removing multiples of each distinct prime divisor removes exactly the numbers that share a nontrivial gcd with \(n\).
The product formula works because phi is multiplicative on coprime arguments.

\section*{Complexity Analysis}

If factorization is done by trial division, computing \(\varphi(n)\) is \(O(\sqrt{n})\). With precomputed SPF or a phi
sieve for many queries, the per-query or all-values cost can be much better.

\section*{Implementation}

The code includes:

\begin{itemize}
  \item \texttt{phi\_single(n)} from factorization,
  \item a linear-style sieve for \(\varphi(1 \dots N)\).
\end{itemize}

\section*{Common Pitfalls}

\begin{itemize}
  \item Updating the result once per prime power instead of once per distinct prime.
  \item Using Euler's theorem where the base is not coprime to the modulus.
  \item Recomputing phi from scratch many times when a sieve would be better.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Phi sieve for all values up to \(N\).
  \item Use phi inside modular-exponent cycles.
  \item General multiplicative-function preprocessing alongside Mobius and divisor counts.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Compute \(\varphi(n)\) for many values.
  \item Count reduced fractions or coprime pairs.
  \item Evaluate large exponents modulo \(m\) with phi-based reasoning.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/algebra/phi-function.html}{cp-algorithms: Euler Totient Function}.
  \item \href{https://usaco.guide/CPH.pdf}{Competitive Programmer's Handbook}.
  \item \href{https://codeforces.com/catalog?locale=en}{Codeforces Catalog}.
\end{itemize}
