\section*{Overview}

The convex hull is the smallest convex polygon containing a set of points. In competitive programming it is often the
first real geometry construction: once the hull is built, diameter, tangents, support lines, and several optimization
problems become easier.

\section*{When to Use It}

Use a convex hull when:

\begin{itemize}
  \item the answer depends only on the extreme points,
  \item interior points are irrelevant noise,
  \item you need a clean polygonal envelope before running rotating calipers or tangent queries.
\end{itemize}

\section*{Core Idea}

The monotone-chain algorithm sorts the points, then builds the lower and upper hulls separately. While scanning, it
removes the last point whenever the last two hull points together with the new point fail the chosen turn condition.

That means the hull is maintained as a sequence of valid turns instead of recomputed from scratch.

\section*{Key Insight}

After sorting by \((x, y)\), every point is processed in the only order that matters for the lower and upper envelopes.
If a point creates a non-left turn on the lower hull, it can never belong to the final lower envelope, so it is safe
to remove immediately.

\section*{Operations / Main Technique}

\begin{itemize}
  \item sort points lexicographically,
  \item build the lower hull,
  \item build the upper hull,
  \item concatenate them while avoiding duplicate endpoints.
\end{itemize}

\paragraph{Worked example.}
If three consecutive candidate points make a clockwise turn while building the lower hull, the middle point lies inside
or on the boundary of the current envelope and can be removed.

\section*{Correctness Intuition}

The hull must be convex, so each chain must keep turning in the same direction. Whenever three consecutive points break
that rule, the middle one cannot be an extreme vertex of that chain. Repeating this local repair produces the global
envelope.

\section*{Complexity Analysis}

Sorting dominates the cost:
\begin{itemize}
  \item time: \(O(n \log n)\),
  \item memory: \(O(n)\).
\end{itemize}

\section*{Implementation}

The code uses monotone chain on integer points and returns the hull in counterclockwise order without repeating the
first vertex at the end.

The only policy choice is how to treat collinear boundary points. The sample keeps the extreme endpoints and removes
intermediate collinear points from each edge.

\section*{Common Pitfalls}

\begin{itemize}
  \item Forgetting to remove duplicate input points first.
  \item Using the wrong turn condition for the chosen collinearity policy.
  \item Mishandling the cases \(n = 0\), \(1\), or \(2\).
  \item Expecting interior collinear points to remain when the implementation intentionally compresses edges.
\end{itemize}

\section*{Variants / Extensions}

\begin{itemize}
  \item Keep all boundary points by changing the pop condition.
  \item Rotating calipers on the hull for diameter or width.
  \item Dynamic hull maintenance, which is a different problem and much harder.
\end{itemize}

\section*{Practice Problems}

\begin{itemize}
  \item Build the hull of a planar point set.
  \item Compute the diameter of a point set after building the hull.
  \item Count or list the vertices of the outer envelope only.
\end{itemize}

\section*{References}

\begin{itemize}
  \item \href{https://cp-algorithms.com/geometry/convex-hull.html}{cp-algorithms: Convex Hull Construction}.
  \item \href{https://usaco.guide/plat/convex-hull}{USACO Guide: Convex Hull}.
  \item \href{https://codeforces.com/catalog}{Codeforces Catalog}.
\end{itemize}
