¥documentclass[11pt]{article}
¥usepackage{amsmath,amssymb,amsthm}
¥usepackage[margin=1in]{geometry}
¥newtheorem{theorem}{Theorem}
¥newtheorem{lemma}[theorem]{Lemma}
¥title{Problem 365: A Huge Binomial Coefficient}
¥author{Project Euler Solutions}
¥date{}
¥begin{document}
¥maketitle

¥section{Problem Statement}

Let $n = 10^{18}$ and $k = 10^{9}$. For each triple of distinct primes $(p_1, p_2, p_3)$ with $1000 < p_i < 5000$, compute $¥binom{n}{k} ¥bmod (p_1 p_2 p_3)$ via the Chinese Remainder Theorem. Find the sum of all such results over all $¥binom{574}{3}$ triples.

¥section{Mathematical Foundation}

¥begin{theorem}[Lucas' Theorem]
Let $p$ be prime. Write $n = ¥sum_{i} n_i p^i$ and $k = ¥sum_{i} k_i p^i$ in base‾$p$. Then
¥[
¥binom{n}{k} ¥equiv ¥prod_{i} ¥binom{n_i}{k_i} ¥pmod{p},
¥]
where $¥binom{n_i}{k_i} = 0$ if $k_i > n_i$.
¥end{theorem}

¥begin{proof}
In $¥mathbb{F}_p[x]$, $(1+x)^p = 1 + x^p$ since $¥binom{p}{j} ¥equiv 0 ¥pmod{p}$ for $1 ¥le j ¥le p-1$. By induction, $(1+x)^{p^i} = 1 + x^{p^i}$. Therefore
¥[
(1+x)^n = ¥prod_i ¥bigl(1 + x^{p^i}¥bigr)^{n_i} ¥pmod{p}.
¥]
Extracting the coefficient of $x^k = x^{¥sum k_i p^i}$ from both sides gives $¥binom{n}{k} ¥equiv ¥prod_i ¥binom{n_i}{k_i}$.
¥end{proof}

¥begin{theorem}[Chinese Remainder Theorem]
Let $m_1, m_2, m_3$ be pairwise coprime. Set $M = m_1 m_2 m_3$ and $M_i = M/m_i$. Given residues $r_i$, the unique solution modulo‾$M$ is
¥[
x ¥equiv ¥sum_{i=1}^{3} r_i ¥cdot M_i ¥cdot (M_i^{-1} ¥bmod m_i) ¥pmod{M}.
¥]
¥end{theorem}

¥begin{proof}
Since $¥gcd(M_i, m_i) = 1$, the inverse $M_i^{-1} ¥bmod m_i$ exists. The term $r_i M_i (M_i^{-1} ¥bmod m_i)$ is congruent to $r_i$ modulo‾$m_i$ and to‾$0$ modulo‾$m_j$ for $j ¥neq i$. Summing gives the solution. Uniqueness follows since $|¥mathbb{Z}/M¥mathbb{Z}| = ¥prod |¥mathbb{Z}/m_i¥mathbb{Z}|$.
¥end{proof}

¥begin{lemma}[Prime Count]
There are exactly $574$ primes in $(1000, 5000)$, giving $¥binom{574}{3} = 31{,}436{,}024$ triples.
¥end{lemma}

¥section{Algorithm}

¥begin{verbatim}
function solve():
    primes = sieve(1000, 5000)   // 574 primes
    for p in primes:
        r[p] = lucas(10^18, 10^9, p)
    S = 0
    for each triple (p1, p2, p3):
        S += CRT(r[p1], r[p2], r[p3], p1, p2, p3)
    return S
¥end{verbatim}

¥section{Complexity Analysis}

¥begin{itemize}
¥item ¥textbf{Time:} $O(¥pi^3)$ where $¥pi = 574$. Each CRT evaluation is $O(1)$; total $¥approx 3.1 ¥times 10^7$ operations.
¥item ¥textbf{Space:} $O(¥pi)$ for the residue array.
¥end{itemize}

¥section{Answer}

¥[
¥boxed{162619462356610313}
¥]

¥end{document}
